---
title: "Filtering dataset and DEA iPSC-DA time course"
author: "J Monzón Sandoval"
date: 'Last update: `r date()`'
output: 
  html_document:
    theme: lumen
    toc: yes
    number_sections: true
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

library(Seurat)
library(scater)
#library(ouija)

library(stringr)
library(ggplot2)
library(ggsci)
library(cowplot)
library(MetBrewer)
library(RColorBrewer)

library(corrgram)
library(venn)
library(reshape)
library(pheatmap)

```

# Gene expression data

```{r GeneExpression}

# Gene_ID	Gene_Name	Gene_Biotype	Gene_Length # Columns in gene info tab 
# ensembl annotations release 99
gene_info <- read.table("./05_MappaAnalyzer_BFP/HET1D21_S2/gene_info.csv", 
                        sep = ",", header = TRUE, stringsAsFactors = FALSE)

# All data matrices
fn <- list.files(path = "./05_MappaAnalyzer_BFP", 
                 pattern = "_genematrix.csv", 
                 recursive = TRUE, full.names = TRUE)
fn <- fn[str_detect(string = fn, pattern = "_incl_introns_", 
                    negate = TRUE)] # change to FALSE to include introns

# Need to combine counts from the different samples
t1 <- read.table(file = fn[1], sep = ",", header = TRUE, row.names = 1)

for (i in c(2:length(fn))){
  ti <- read.table(file = fn[i], sep = ",", header = TRUE, row.names = 1)
  t1 <- cbind(t1, ti)
}

ged <- t1
rm(i, t1, ti, fn)

ged <- as.matrix(ged)

```

```{r PhenoPart1}

# Phenotype data 

# Mappa Stats
pheno <- read.table("./06_Pheno/Pheno_MappaStats_L1234.txt", sep = "\t", stringsAsFactors = FALSE, header = TRUE)

# Phenotype
Clone <- vector()
Clone[str_detect(pheno$Sample, "HET1")] <- "L25"
Clone[str_detect(pheno$Sample, "HET2")] <- "L35"

Day <- vector()
Day[str_detect(pheno$Sample, "D21")] <- "D21"
Day[str_detect(pheno$Sample, "D30")] <- "D30"
Day[str_detect(pheno$Sample, "D45")] <- "D45"
Day[str_detect(pheno$Sample, "D65")] <- "D65"

Control <- vector()
Control[str_detect(pheno$Sample, "Neg_Ctrl")] <- "Neg_Ctrl"
Control[str_detect(pheno$Sample, "Pos_Ctrl")] <- "Pos_Ctrl"

Protocol <- rep("WHOLE", length = nrow(pheno))
Protocol[str_detect(pheno$Sample, "NUCLEI")] <- "NUCLEI"
Protocol[str_detect(pheno$Sample, "Neg_Ctrl")] <- "Neg_Ctrl"
Protocol[str_detect(pheno$Sample, "Pos_Ctrl")] <- "Pos_Ctrl"

pheno <- data.frame(pheno[, 1:2], 
                    Clone  = Clone, 
                    Day = Day,
                    Control = Control,
                    Protocol = Protocol, 
                    Percent_Mapped = pheno$Mapped_Reads * 100 / pheno$Barcoded_Reads, 
                    Percent_Multimapped_Reads = pheno$Multimapped_Reads * 100 / pheno$Mapped_Reads,
                    Percent_Uniquely_Mapped = pheno$Uniquely_Mapped_Reads * 100 / pheno$Mapped_Reads,
                    Percent_Exon_Reads = pheno$Exon_Reads * 100 / pheno$Uniquely_Mapped,
                    Percent_Ambiguous_Exon_Reads = pheno$Ambiguous_Exon_Reads * 100 / pheno$Uniquely_Mapped,
                    Percent_Intron_Reads = pheno$Intron_Reads * 100 / pheno$Uniquely_Mapped,
                    Percent_Ambiguous_Intron_Reads = pheno$Ambiguous_Intron_Reads * 100 / pheno$Uniquely_Mapped,
                    Percent_Intergenic_Reads = pheno$Intergenic_Reads * 100 / pheno$Uniquely_Mapped,
                    Percent_Mitochondrial_Reads = pheno$Mitochondrial_Reads * 100 / pheno$Uniquely_Mapped,
                    Percent_Ribosomal_Reads = pheno$Ribosomal_Reads * 100 / pheno$Uniquely_Mapped,
                    pheno[, 3:ncol(pheno)])

rm(Day, Control, Clone, Protocol)

```

```{r PhenoPart2}

# Feature annotation and counts only for Protein Coding genes
fd <- gene_info[which(gene_info$Gene_Biotype == "protein_coding"), ]
rownames(fd) <- fd$Gene_ID
pc_genes <- rownames(fd)
pc_counts <- ged[pc_genes, ]

# Rearrange order of gene expression data to pheno
ged <- ged[, pheno$Barcode]

# single cell object
sce <- SingleCellExperiment(assays = list(counts = pc_counts),
                            colData = pheno)

rowData(sce) <- fd

# Calculate QC for cells
QCcells <- perCellQCMetrics(sce, percent_top = c(10, 20, 50, 100)) 

# Normalize data 
cpm(sce) <- calculateCPM(sce)
tpm(sce) <- calculateTPM(sce, lengths = rowData(sce)[, "Gene_Length"])

# Add QC for cells and number of genes detected to colData
genes_detected_1TPM <- apply(X = tpm(sce), MARGIN = 2, FUN = function(x) length(which(x >= 1)))
genes_detected_5TPM <- apply(X = tpm(sce), MARGIN = 2, FUN = function(x) length(which(x >= 5)))
genes_detected_1cnt <- apply(X = counts(sce), MARGIN = 2, FUN = function(x) length(which(x >= 1)))
genes_detected_5cnt <- apply(X = counts(sce), MARGIN = 2, FUN = function(x) length(which(x >= 5)))

colData(sce) <- cbind(colData(sce), QCcells)
colData(sce) <- cbind(colData(sce), data.frame(genes_detected_1TPM = genes_detected_1TPM,
                                               genes_detected_5TPM = genes_detected_5TPM, 
                                               genes_detected_1cnt = genes_detected_1cnt,
                                               genes_detected_5cnt = genes_detected_5cnt))


rm(fd, QCcells)
rm(genes_detected_1TPM, genes_detected_5TPM, genes_detected_1cnt, genes_detected_5cnt)

# Cell pheno data
pheno <- as.data.frame(colData(sce), row.names = sce$Barcode)

rm(sce)

```

+ Total features: `r nrow(ged)`.

+ Total number of barcodes `r ncol(ged)`.

# Data filtering

## Protein coding genes

```{r}

# Seurat object
iDA <- CreateSeuratObject(counts = pc_counts, meta.data = pheno, project = "TimeCourse", min.cells = 0, min.features = 0)

rm(ged, pc_counts, pc_genes)

```

+ Total features: `r nrow(iDA)` .

+ Total number of barcodes `r ncol(iDA)`.

## Filtering cells 

### Cells with low number of genes detected.

```{r fig.height=3, fig.width=4}

mingenes <- 2000

plot(density(pheno$genes_detected_1cnt), main = "Genes detected (1 count)")
abline( v = mingenes, col = "red", lty = 2)
text(x = mingenes, y = 0.00010, labels = paste(mingenes, "genes"), srt = 90, pos = 2)

```

```{r}

cat("\n + Barcodes per sample: \n")
table(pheno$Sample)
cat("\n")

cat("\n + Barcodes per sample with < ", mingenes, "genes detected: \n")
table(pheno$Sample[which(pheno$genes_detected_1cnt < mingenes)])
cat("\n")

cat("\n + Barcodes per sample with > ", mingenes, "genes detected: \n")
table(pheno$Sample[which(pheno$genes_detected_1cnt > mingenes)])
cat("\n")

iDA <- subset(x = iDA, cells = pheno$Barcode[which(pheno$genes_detected_1cnt > mingenes)])

rm(mingenes)

```

+ Total features: `r nrow(iDA)` .

+ Total number of barcodes `r ncol(iDA)`.

### Internal controls

```{r}

cat("+ We furhter removed", length(which(iDA$Sample == "Neg_Ctrl")), " Neg_Ctrl samples \n")
cat("+ We furhter removed", length(which(iDA$Sample == "Pos_Ctrl")), " Pos_Ctrl samples \n")

iDA <- subset(x = iDA, cells = iDA$Barcode[!(iDA$Sample == "Neg_Ctrl")])
iDA <- subset(x = iDA, cells = iDA$Barcode[!(iDA$Sample == "Pos_Ctrl")])

```

+ Total features: `r nrow(iDA)` .

+ Total number of barcodes `r ncol(iDA)`.

### Single nuclei at D45

```{r}

cat("+ We furhter removed", length(which(iDA$Sample == "HET1D45NUCLEI")), " HET1D45NUCLEI samples \n")

iDA <- subset(x = iDA, cells = iDA$Barcode[!(iDA$Sample == "HET1D45NUCLEI")])

```


### Cells with high % of mitochondrial reads

```{r}

cat("+ We furhter removed", length(which(iDA$Percent_Mitochondrial_Reads > 20)), 
    "cell with high % of mitochondrial reads (> 20%) \n")

iDA <- subset(x = iDA, cells = iDA$Barcode[!(iDA$Percent_Mitochondrial_Reads > 20)])

cat("At the end", sum(table(iDA$Sample)), "cells remained:\t")
table(iDA$Sample)

```

+ Total features: `r nrow(iDA)` .

+ Total number of barcodes `r ncol(iDA)`.


## Filtering genes

## Genes consistently expressed across cells

```{r warning=FALSE}

cells_nonzero <- apply(X = iDA[["RNA"]]@counts, MARGIN = 1, FUN = function(x) length(which(x != 0)))
cells_above1 <- apply(X = iDA[["RNA"]]@counts, MARGIN = 1, FUN = function(x) length(which(x > 1)))
cells_above5 <- apply(X = iDA[["RNA"]]@counts, MARGIN = 1, FUN = function(x) length(which(x > 5)))
cells_above10 <- apply(X = iDA[["RNA"]]@counts, MARGIN = 1, FUN = function(x) length(which(x > 10)))

output <- matrix(NA, nrow = 4, ncol = 100, 
                 dimnames = list(c("Non zero counts", "More than 1 count", 
                                   "More than 5 counts", "More than 10 counts"), 
                                 seq(.01, 1, .01)))

percent_cells <- seq(.01, 1, .01)

for(i in c(1:100)){
  output[1, i] <- length(which(cells_nonzero > ncol(iDA) * percent_cells[i]))
  output[2, i] <- length(which(cells_above1 > ncol(iDA) * percent_cells[i]))
  output[3, i] <- length(which(cells_above5 > ncol(iDA) * percent_cells[i]))
  output[4, i] <- length(which(cells_above10 > ncol(iDA) * percent_cells[i]))
  
}

rm(i, percent_cells)

output <- melt(output)
colnames(output) <- c("Counts", "Fractions_Of_cells", "Genes")

```


```{r fig.width=5.5, fig.height=3.5}

g1 <- ggplot(data = output, mapping = aes(x = Fractions_Of_cells, y = Genes, color = Counts)) + 
  geom_point(size = 1) + 
  scale_color_lancet() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) 

g1 

rm(g1, output)

```

## Filter lowly expressed genes

+ Keep genes with more than 1 count in at least 5 percent of the filtered cells.

```{r}

#keep_genes <- rownames(iDA)[which(cells_nonzero >= (ncol(iDA) * .1))]
keep_genes <- rownames(iDA)[which(cells_above1 >= (ncol(iDA) * .05))]
iDA <- subset(x = iDA, features = keep_genes)

rm(cells_above1, cells_above10, cells_above5, cells_nonzero)
rm(keep_genes)

```

+ Total features: `r nrow(iDA)` .

+ Total number of barcodes `r ncol(iDA)`.


# Analysis 

## Normalization and clustering 

+ Carried out in all filtered cells.

+ Normalization LogNormalize (scale factor 1e6).

+ PCA based on the top 5000 most variable features.

+ UMAP based on top 10 PCS.


```{r fig.height=3, fig.width=4}

iDA <- NormalizeData(iDA, normalization.method = "LogNormalize", scale.factor = 1e6)
iDA <- FindVariableFeatures(iDA, selection.method = "vst", nfeatures = 5000)
all_genes <- rownames(iDA)

# Scale on all data run PC in top 5000 most variable features
iDA <- ScaleData(iDA, features = all_genes)
iDA <- RunPCA(iDA, features = VariableFeatures(object = iDA), verbose = FALSE)
# Elbow plot
barplot(iDA@reductions$pca@stdev[1:20], xlab = "PC", ylab = "Standard Deviation", 
        names.arg = 1:20, las = 2, col = grey.colors(20))

# Cell cycle score
s_genes <- gene_info$Gene_ID[match(cc.genes$s.genes, table = gene_info$Gene_Name)]
g2m_genes <- gene_info$Gene_ID[match(cc.genes$g2m.genes, table = gene_info$Gene_Name)]
s_genes <- intersect(rownames(iDA), s_genes)
g2m_genes <- intersect(rownames(iDA), g2m_genes)
iDA <- CellCycleScoring(iDA, s.features = s_genes, g2m.features = g2m_genes, set.ident = FALSE)
rm(g2m_genes, s_genes)

# UMAP and cell clustering 
iDA <- RunUMAP(iDA, reduction = "pca", dims = 1:10)
iDA <- FindNeighbors(iDA, dims = 1:10)
iDA <- FindClusters(iDA, resolution = 0.5)

# Threshold for DEGs
cut_adjp <- 0.05

```

## Differential expression by Day

+ Differential expression analysis (DEA) based on Wilcox test (lfc > 0.25 and adjusted p value 0.05).


```{r DEGsDay}

# Differential expression analysis by Day

Idents(object = iDA) <- iDA$Day

DEA_Day30_Day21 <- FindMarkers(object = iDA, ident.1 = "D30", ident.2 = "D21",
                               logfc.threshold = 0.25, test.use = "wilcox")

DEA_Day45_Day30 <- FindMarkers(object = iDA, ident.1 = "D45", ident.2 = "D30",
                               logfc.threshold = 0.25, test.use = "wilcox")

DEA_Day65_Day45 <- FindMarkers(object = iDA, ident.1 = "D65", ident.2 = "D45",
                               logfc.threshold = 0.25, test.use = "wilcox")

DEA_Day30_Day21 <- DEA_Day30_Day21[which(DEA_Day30_Day21$p_val_adj < cut_adjp), ]
DEA_Day45_Day30 <- DEA_Day45_Day30[which(DEA_Day45_Day30$p_val_adj < cut_adjp), ]
DEA_Day65_Day45 <- DEA_Day65_Day45[which(DEA_Day65_Day45$p_val_adj < cut_adjp), ]
  
# Number of DEGs by Day
nDEG_Day <- c(nrow(DEA_Day30_Day21), nrow(DEA_Day45_Day30), nrow(DEA_Day65_Day45))
nDEG_Day <- data.frame(Comparison = c("D30_vs_D21", "D45_vs_D30", "D65_vs_D45"), DEG = nDEG_Day)

# Add gene names to DEA 
DEA_Day30_Day21 <- data.frame(DEA_Day30_Day21, Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_Day30_Day21), table = gene_info$Gene_ID)])
DEA_Day45_Day30 <- data.frame(DEA_Day45_Day30, Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_Day45_Day30), table = gene_info$Gene_ID)])
DEA_Day65_Day45 <- data.frame(DEA_Day65_Day45, Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_Day65_Day45), table = gene_info$Gene_ID)])

```

## Differential expression by Clone

+ Differential expression analysis (DEA) based on Wilcox test (lfc > 0.25 and adjusted p value 0.05).

```{r DEAClone}

# Differentially expressed by Clone 
Idents(object = iDA) <- iDA$Clone

DEA_Clone <- FindMarkers(object = iDA, ident.1 = "L25", ident.2 = "L35", 
                         logfc.threshold = 0.25, test.use = "wilcox")
DEA_Clone <- DEA_Clone[which(DEA_Clone$p_val_adj < cut_adjp), ]


# Number of DEG by Clone
nDEG_Clone <- c(nrow(DEA_Clone))

nDEG_Clone <- data.frame(Comparison = "L35_vs_L25",
                         DEG = nDEG_Clone)

# Add gene names to DEA 
DEA_Clone <- data.frame(DEA_Clone, Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_Clone), table = gene_info$Gene_ID)])

```

## Cluster gene markers

+ Gene markers per cluster (lfc > 0.25, adjusted p value < 0.05 and only positive true).

```{r DEACluster}

# Differential expression analysis: cluster markers # 

Idents(object = iDA) <- iDA$seurat_clusters

DEA_c0 <- FindMarkers(object = iDA, ident.1 = "0",
                      logfc.threshold = 0.25, test.use = "wilcox", 
                      only.pos = TRUE)

DEA_c1 <- FindMarkers(object = iDA, ident.1 = "1",
                      logfc.threshold = 0.25, test.use = "wilcox", 
                      only.pos = TRUE)

DEA_c2 <- FindMarkers(object = iDA, ident.1 = "2",
                      logfc.threshold = 0.25, test.use = "wilcox", 
                      only.pos = TRUE)

DEA_c3 <- FindMarkers(object = iDA, ident.1 = "3",
                      logfc.threshold = 0.25, test.use = "wilcox", 
                      only.pos = TRUE)

DEA_c4 <- FindMarkers(object = iDA, ident.1 = "4",
                      logfc.threshold = 0.25, test.use = "wilcox", 
                      only.pos = TRUE)

DEA_c5 <- FindMarkers(object = iDA, ident.1 = "5",
                      logfc.threshold = 0.25, test.use = "wilcox", 
                      only.pos = TRUE)

DEA_c6 <- FindMarkers(object = iDA, ident.1 = "6",
                      logfc.threshold = 0.25, test.use = "wilcox", 
                      only.pos = TRUE)

DEA_c0 <- DEA_c0[which(DEA_c0$p_val_adj < cut_adjp), ]
DEA_c1 <- DEA_c1[which(DEA_c1$p_val_adj < cut_adjp), ]
DEA_c2 <- DEA_c2[which(DEA_c2$p_val_adj < cut_adjp), ]
DEA_c3 <- DEA_c3[which(DEA_c3$p_val_adj < cut_adjp), ]
DEA_c4 <- DEA_c4[which(DEA_c4$p_val_adj < cut_adjp), ]
DEA_c5 <- DEA_c5[which(DEA_c5$p_val_adj < cut_adjp), ]
DEA_c6 <- DEA_c6[which(DEA_c6$p_val_adj < cut_adjp), ]

# Add gene names to DEA
DEA_c0 <- data.frame(DEA_c0, Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_c0), table = gene_info$Gene_ID)])
DEA_c1 <- data.frame(DEA_c1, Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_c1), table = gene_info$Gene_ID)])
DEA_c2 <- data.frame(DEA_c2, Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_c2), table = gene_info$Gene_ID)])
DEA_c3 <- data.frame(DEA_c3, Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_c3), table = gene_info$Gene_ID)])
DEA_c4 <- data.frame(DEA_c4, Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_c4), table = gene_info$Gene_ID)])
DEA_c5 <- data.frame(DEA_c5, Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_c5), table = gene_info$Gene_ID)])
DEA_c6 <- data.frame(DEA_c6, Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_c6), table = gene_info$Gene_ID)])

# Number of DEGs per Cluster
nDEG_Cluster <- c(nrow(DEA_c0), nrow(DEA_c1), nrow(DEA_c2),
                  nrow(DEA_c3), nrow(DEA_c4), nrow(DEA_c5),
                  nrow(DEA_c6))

nDEG_Cluster <- data.frame(Cluster = as.character(seq(0, 6)), #
                           DEG = nDEG_Cluster)

# Top DEGs by Cluster
DEGs_Cluster_top <- unique(c(head(rownames(DEA_c0), 10), head(rownames(DEA_c1), 10),
                             head(rownames(DEA_c2), 10), head(rownames(DEA_c3), 10),
                             head(rownames(DEA_c4), 10), head(rownames(DEA_c5), 10),
                             head(rownames(DEA_c6), 10)))


```

## Known cell type markers

```{r genemarkers}

# Cell type markers
gm <- read.table("./06_Pheno/CellType_Markers_Subset2.txt", sep = "\t", header = TRUE)
gm <- data.frame(gm, Gene_ID = gene_info$Gene_ID[match(gm$Gene_Name, gene_info$Gene_Name)])

cat("Markers not found in filtered data:", gm$Gene_Name[!(gm$Gene_ID %in% all_genes)],"\n")

gm <- gm[gm$Gene_ID %in% all_genes, ]
gm$Gene_ID <- factor(gm$Gene_ID, levels = gm$Gene_ID) 

```

## Overlap cluster markers with cell type markers

```{r celltypeannot, results='asis'}

# Overlap of cluster markers with cell type markers
cat("\n\n\ + Overlap with cluster 0 markers: \n")
gm[gm$Gene_ID %in% rownames(DEA_c0),]
cat("\n\n\ + Overlap with cluster 1 markers: \n")
gm[gm$Gene_ID %in% rownames(DEA_c1),]
cat("\n\n\ + Overlap with cluster 2 markers: \n")
gm[gm$Gene_ID %in% rownames(DEA_c2),]
cat("\n\n\ + Overlap with cluster 3 markers: \n")
gm[gm$Gene_ID %in% rownames(DEA_c3),]
cat("\n\n\ + Overlap with cluster 4 markers: \n")
gm[gm$Gene_ID %in% rownames(DEA_c4),]
cat("\n\n\ + Overlap with cluster 5 markers: \n")
gm[gm$Gene_ID %in% rownames(DEA_c5),]
cat("\n\n\ + Overlap with cluster 6 markers: \n")
gm[gm$Gene_ID %in% rownames(DEA_c6),]

```

## human Embryonic Midbrain (hEM)

+ Data obtained from GEO [GSE76381](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE76381).

```{r hmref}

# Human midbrain dataset (for reference)
# Pheno data for cell, including cell type and time points
hm_pheno <- read.table("../iPSC_DA_comparison/04_ExternalData/RNAseq_data/GSE76381_EmbryoMoleculeCounts.cef.txt", 
                            header = FALSE, sep = "\t", nrows = 3, stringsAsFactors = TRUE, skip = 1)

hm_pheno <- hm_pheno[, -1]
rownames(hm_pheno) <- hm_pheno$V2
hm_pheno <- hm_pheno[, -1]
hm_pheno <- t(hm_pheno)
hm_pheno <- as.data.frame(hm_pheno)

hm_celltypes <- read.table("../iPSC_DA_comparison/04_ExternalData/RNAseq_data/GSE76381_EmbryoMoleculeCounts.CelltypeDescription.txt", 
                           sep = "\t", header = TRUE)

hm_pheno <- data.frame(hm_pheno, 
                       cell_type = hm_celltypes$cell_type_lv1[match(hm_pheno$Cell_type,
                                                                        table = hm_celltypes$cell_type)],
                       row.names = hm_pheno$Cell_ID)

# table(colnames(hm_counts) == rownames(hm_pheno))


# La Manno 2016 counts
hm_counts <- read.table("../iPSC_DA_comparison/04_ExternalData/RNAseq_data/GSE76381_EmbryoMoleculeCounts.cef.txt", 
                            header = FALSE, sep = "\t", skip = 5, row.names = 1)
hm_counts <- hm_counts[, -1] # Empty column 
colnames(hm_counts) <- hm_pheno$Cell_ID

hm <- CreateSeuratObject(counts = hm_counts, meta.data = hm_pheno, 
                         project = "HumanMidbrain", min.cells = 0, min.features = 0)

rm(hm_counts, hm_pheno, hm_celltypes)

```

+ Subset progenitors, neuroblast and dopaminergic neurons.

+ Use common features with our dataset for integration.

```{r hmfilt}

# Filter out endothelial, pericytes, microglia, OMTN, red nucleus, gabaergic, serotonergic, OPC, radial glia, and unknown cells from hm
hm <- subset(x = hm, cells = hm$Cell_ID[hm$cell_type != "endothelial"])
hm <- subset(x = hm, cells = hm$Cell_ID[hm$cell_type != "pericytes"])
hm <- subset(x = hm, cells = hm$Cell_ID[hm$cell_type != "microglia"])
hm <- subset(x = hm, cells = hm$Cell_ID[hm$cell_type != "OMTN"])
hm <- subset(x = hm, cells = hm$Cell_ID[hm$cell_type != "red nucleus"])
hm <- subset(x = hm, cells = hm$Cell_ID[hm$cell_type != "gabaergic"])
hm <- subset(x = hm, cells = hm$Cell_ID[hm$cell_type != "serotonergic"])
hm <- subset(x = hm, cells = hm$Cell_ID[hm$cell_type != "OPC"])
hm <- subset(x = hm, cells = hm$Cell_ID[hm$cell_type != "unknown"])
hm <- subset(x = hm, cells = hm$Cell_ID[hm$cell_type != "radial glia"])

# Common background set of genes shared between datasets
hm_bg <- rownames(hm)

tomap  <- gene_info[gene_info$Gene_ID %in% rownames(iDA), ]
rownames(tomap) <- 1:nrow(tomap)
if(length(which(duplicated(tomap$Gene_Name) == TRUE)) > 0 ) {
  dps <- tomap$Gene_Name[which(duplicated(x = tomap$Gene_Name) == TRUE)]
  tomap <- tomap[(!tomap$Gene_Name %in% dps), ]
  rm(dps)
}

bg <- intersect(hm_bg, tomap$Gene_Name) # # length(bg) # 12309
bg_ens <- tomap$Gene_ID[match(bg, table = tomap$Gene_Name)]

# subset Human midbrain datasets
hm <- subset(x = hm, features = bg)
hm_counts <- hm@assays$RNA@counts
rownames(hm_counts) <- tomap$Gene_ID[match(x = rownames(hm_counts), table = tomap$Gene_Name)]
hm_pheno <- hm@meta.data
rm(hm)

hm <- CreateSeuratObject(counts = hm_counts, meta.data = hm_pheno,
                         project = "HumanMidbrain", min.cells = 0, min.features = 0)
rm(hm_counts, hm_pheno)

# subset counts from iPSC dataset
iDAs <- subset(iDA, features = bg_ens)
iDAs_counts <- iDAs@assays$RNA@counts
iDAs_pheno <- iDAs@meta.data
rm(iDAs)

iDAs <- CreateSeuratObject(counts = iDAs_counts, meta.data = iDAs_pheno,
                         project = "TimeCourse", min.cells = 0, min.features = 0)
rm(iDAs_counts, iDAs_pheno)

# Seurat Object list of both human midbrain and iPSC differentiation into DA neurons
so_list <- list(HumanMidbrain = hm, iPSCTimeCourse = iDAs)
rm(hm, iDAs)

rm(hm_bg, bg, bg_ens, tomap)

```

+ Annotate our iPSC to DA differentiation time course data using the hEM as reference.

```{r hmlabelt}

# Label transfer from neuroblast, neuronal progenitors and Dopaminergic neurons

for (i in 1:length(so_list)) {
    so_list[[i]] <- NormalizeData(so_list[[i]], normalization.method = "LogNormalize", scale.factor = 1e6)
    so_list[[i]] <- FindVariableFeatures(so_list[[i]], selection.method = "vst", nfeatures = 5000, verbose = FALSE)
    so_list[[i]] <- ScaleData(so_list[[i]], features = all_genes)
    so_list[[i]] <- RunPCA(so_list[[i]], features = VariableFeatures(object = so_list[[i]]), verbose = FALSE)
    so_list[[i]] <- RunUMAP(so_list[[i]], reduction = "pca", dims = 1:10)
}
rm(i)

#data_anchors <- FindIntegrationAnchors(object.list = so_list, dims = 1:10)
#integrated_data <- IntegrateData(anchorset = data_anchors, new.assay.name = "both")

tranf_anchors <- FindTransferAnchors(reference = so_list[["HumanMidbrain"]], 
                                     query = so_list[["iPSCTimeCourse"]], 
                                     dims = 1:10, reference.reduction = "pca")

predictions <- TransferData(anchorset = tranf_anchors, 
                            refdata = so_list[["HumanMidbrain"]]$cell_type, 
                            dims = 1:10)

iDA <- AddMetaData(object = iDA, 
                   metadata = predictions$predicted.id, 
                   col.name = "Predicted")
iDA <- AddMetaData(object = iDA, 
                   metadata = predictions$prediction.score.dopaminergic, 
                   col.name = "PS_dopaminergic")
iDA <- AddMetaData(object = iDA, metadata = predictions$prediction.score.neuronal.progenitor, 
                   col.name = "PS_neuronal_progenitor")
iDA <- AddMetaData(object = iDA, metadata = predictions$prediction.score.neuroblasts, 
                   col.name = "PS_neuroblasts")
iDA <- AddMetaData(object = iDA, 
                   metadata = predictions$prediction.score.max, 
                   col.name = "Max_Prediction_Score")

#iDA <- AddMetaData(object = iDA, metadata = predictions$prediction.score.radial.glia, col.name = "Score_radial_glia")

rm(tranf_anchors, predictions)

```

+ Project our dataset into the PCs of the hEM.

```{r projec}

# Cell loadings from human midbrain
hm_loads <- so_list$HumanMidbrain@reductions$pca@feature.loadings[, 1:10]

# Gene expression form iPSC derived cells
iDA_ge <- as.matrix(so_list$iPSCTimeCourse@assays$RNA@data)
iDA_ge <- iDA_ge[rownames(hm_loads), ]

# New Projected PCA for iPSC data
ipsc_projected <- scale(t(iDA_ge), center = TRUE, scale = TRUE) %*% hm_loads

# Data to plot projection into hm 
hm_pcs <- data.frame(so_list$HumanMidbrain@reductions$pca@cell.embeddings,
                             so_list$HumanMidbrain@meta.data)
iDA_pc_proj <- data.frame(ipsc_projected, so_list$iPSCTimeCourse@meta.data)

rm(hm_loads, iDA_ge, ipsc_projected)

```

## Cell type assignment

```{r celltype}

# Assign cell clusters to cell types
cell_type <- as.character(iDA$seurat_clusters)
cell_type[which(cell_type == 0)] <- "DA"
cell_type[which(cell_type == 1)] <- "NB_DA"
cell_type[which(cell_type == 2)] <- "ProgProlif1"
cell_type[which(cell_type == 3)] <- "ProgProlif2"
cell_type[which(cell_type == 4)] <- "Prog1"
cell_type[which(cell_type == 5)] <- "Prog2"
cell_type[which(cell_type == 6)] <- "Astro"

cell_type <- factor(x = cell_type, 
                    levels = c("Prog1", "Prog2", 
                               "ProgProlif1", "ProgProlif2",
                               "NB_DA", "DA", "Astro"))

iDA <- AddMetaData(object = iDA, metadata = cell_type, col.name = "Cell_type")
rm(cell_type)

```

## Differential expression D45 and D65 in DA neurons

```{r}

Idents(iDA) <- iDA$Cell_type

# Focus on DA neuronal cluster
iDA_DA <- subset(iDA, cells = colnames(iDA)[which(iDA$Cell_type == "DA")])

# DA cells per time point
# D30 D45 D65 
#  59  39  62 

# Comparison between later differentiation time points (D65 vs D45).
Idents(iDA_DA) <- iDA_DA$Day
DEA_DA_Day65_Day45 <- FindMarkers(object = iDA_DA, ident.1 = "D65", ident.2 = "D45",
                                  logfc.threshold = 0.25, test.use = "wilcox")
DEA_DA_Day65_Day45 <- DEA_DA_Day65_Day45[which(DEA_DA_Day65_Day45$p_val_adj < cut_adjp), ]
DEA_DA_Day65_Day45 <- data.frame(DEA_DA_Day65_Day45, Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_DA_Day65_Day45), table = gene_info$Gene_ID)])

```

## Code for figures

```{r obj4plots}

# Number of DEGs by Clone and Day for Fig. S3A
nDEG <- rbind(nDEG_Clone, nDEG_Day)

# Updated pheno data for Figs # Add UMAP dims
pheno <- data.frame(iDA@meta.data, iDA@reductions$umap@cell.embeddings[, 1:2])

# Gene markers of gene clusters
DEGs_Cluster <- list(rownames(DEA_c0), rownames(DEA_c1), rownames(DEA_c2),
                     rownames(DEA_c3), rownames(DEA_c4), rownames(DEA_c5), 
                     rownames(DEA_c6))

# Overlap cluster markers and gene markers
ov_gm_DEGclust <- gm[gm$Gene_ID %in% unique(unlist(DEGs_Cluster)), ]

```

```{r OVClusterMarkers}

# Overlap between gene markers
#ov <- matrix(NA, nrow = length(DEGs_Cluster), ncol = length(DEGs_Cluster), 
#             dimnames = list(paste("Cluster", seq(0, length(DEGs_Cluster)-1), sep = "_"),
#                             paste("Cluster", seq(0, length(DEGs_Cluster)-1), sep = "_")))

#for(i in c(1:length(DEGs_Cluster))){
#  ci <- DEGs_Cluster[[i]]
#  for(j in c(1:length(DEGs_Cluster))){
#    cj <- DEGs_Cluster[[j]]
#    ov[i, j] <- length(intersect(ci, cj)) / length(union(ci, cj))
#  }
#}

#ov_anot <- data.frame(Cluster = paste("Cluster", seq(0, length(DEGs_Cluster)-1), sep = "_"),
#                      row.names = paste("Cluster", seq(0, length(DEGs_Cluster)-1), sep = "_"))

#ov_pal <- unlist(as.list(met.brewer(name = "Signac", n = length(DEGs_Cluster))))
#ov_anot_col <- list(Cluster = c("Cluster_0" = ov_pal[1],
#                                "Cluster_1" = ov_pal[2],
#                                "Cluster_2" = ov_pal[3],
#                                "Cluster_3" = ov_pal[4],
#                                "Cluster_4" = ov_pal[5],
#                                "Cluster_5" = ov_pal[6],
#                                "Cluster_6" = ov_pal[7]))

#pheatmap(ov, color = colorRampPalette(brewer.pal(n = 7, name = "Greys"))(100),
#         annotation_colors = ov_anot_col, 
#         annotation_col = ov_anot, 
#         annotation_row = ov_anot, 
#         treeheight_row = 15, 
#         treeheight_col = 15, 
#         border_color = "white", 
#         display_numbers = TRUE, 
#         number_format = "%.2f", 
#         show_colnames = FALSE, 
#         show_rownames = FALSE, 
#         annotation_legend = FALSE)

```

```{r Fig3}

# For cell type annotation of clusters
cls <- aggregate.data.frame(x = pheno[c("UMAP_1", "UMAP_2")], 
                            by = list(pheno$Cell_type), 
                            FUN = median)

# UMAP coloured by cluster
Fig3a <- ggplot(data = pheno, mapping = aes(x = UMAP_1, y = UMAP_2, color = seurat_clusters)) +
  geom_point(size = .3) + 
  scale_color_met_d(name = "Signac") + 
  theme_cowplot() + 
  theme(axis.title = element_text(size = 10),
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        legend.text = element_text(size = 8),
        legend.position = "top", 
        legend.justification = "center", 
        legend.direction = "horizontal", 
        legend.title = element_blank()) +
  ggtitle(label = "") + 
  xlab(label = "UMAP 1") + 
  ylab(label = "UMAP 2") + 
  guides(colour = guide_legend(override.aes = list(size = 2))) 
Fig3a <- Fig3a + annotate("text", x = cls$UMAP_1, y = cls$UMAP_2, label = cls$Group.1)


# UMAP coloured by Day
Fig3b <- ggplot(data = pheno, mapping = aes(x = UMAP_1, y = UMAP_2, color = Day)) + 
  geom_point(size = .3) +  
  scale_color_startrek() + 
  theme_cowplot() + 
  theme(axis.title = element_text(size = 10),
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        legend.text = element_text(size = 8),
        legend.position = "top", 
        legend.justification = "center", 
        legend.direction = "horizontal", 
        legend.title = element_blank()) +
  ggtitle(label = "") + 
  xlab(label = "UMAP 1") + 
  ylab(label = "UMAP 2") + 
  guides(colour = guide_legend(override.aes = list(size = 2), nrow = 2, byrow = TRUE))

# Cells per cluster coloured by Day
Fig3c <- ggplot(data = pheno, mapping = aes(x = seurat_clusters, fill = Day)) + 
  geom_bar() + 
  theme_cowplot() +
  scale_fill_startrek() + 
  xlab("Cluster") + 
  theme(axis.title = element_text(size = 10),
        axis.text = element_text(size = 8), 
        legend.text = element_text(size = 8),
        legend.position = "top", 
        legend.justification = "center", 
        legend.direction = "horizontal", 
        legend.title = element_blank()) +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE))


# Overlapping gene makers with cell types
Fig3d <- DotPlot(iDA, features = ov_gm_DEGclust$Gene_ID, cluster.idents = TRUE) 
Fig3d$data$features.plot <- factor(gene_info$Gene_Name[match(x = Fig3d$data$features.plot, 
                                                             table = gene_info$Gene_ID)],
                                   levels = ov_gm_DEGclust$Gene_Name)
Fig3d <- Fig3d + theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1, size = 8),
                       legend.text = element_text(size = 8),
                       legend.position = "top",
                       legend.justification = "center",
                       legend.direction = "horizontal")


# PC1 in human midbrain 
Fig3e <- ggplot(data = hm_pcs, mapping = aes(x = PC_1, y = cell_type, fill = cell_type)) + 
  geom_violin() + 
  theme_cowplot() + 
  scale_fill_jco() + 
  xlim(-25, 35) + 
  xlab(label = "PC 1") +
  theme(axis.title = element_text(size = 10),
        axis.text = element_text(size = 10),
        legend.text = element_text(size = 8),
        legend.position = "top", 
        legend.justification = "center", 
        legend.direction = "horizontal", 
        legend.title = element_blank()) +  
  geom_vline(xintercept = 0, linetype = "dashed") + 
  guides(fill = guide_legend(nrow = 2, byrow = TRUE))
  
# Projection into PC1 grouped by day
Fig3f <- ggplot(data = iDA_pc_proj, mapping = aes(x = PC_1, y = Day, fill = Day)) + 
  geom_violin() + 
  theme_cowplot() + 
  scale_fill_startrek() + 
  xlim(-25, 35) + 
  xlab(label = "PC 1") +
  theme(axis.title = element_text(size = 10),
        axis.text = element_text(size = 10),
        legend.text = element_text(size = 8),
        legend.position = "top", 
        legend.justification = "center", 
        legend.direction = "horizontal", 
        legend.title = element_blank()) + 
  geom_vline(xintercept = 0, linetype = "dashed") +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE))

# Projection into PC1 grouped by seurat clusters
Fig3g <- ggplot(data = iDA_pc_proj, mapping = aes(x = PC_1, y = seurat_clusters, fill = seurat_clusters)) + 
  geom_violin() + 
  theme_cowplot() + 
  scale_fill_met_d(name = "Signac") + 
  xlim(-25, 35) + 
  xlab(label = "PC 1") +
  theme(axis.title = element_text(size = 10),
        axis.text = element_text(size = 10),
        legend.text = element_text(size = 8),
        legend.position = "top", 
        legend.justification = "center", 
        legend.direction = "horizontal", 
        legend.title = element_blank()) +
  geom_vline(xintercept = 0, linetype = "dashed")

```

```{r FigS3}

# A) Number of differentially expressed genes by Clone and Day
FigS3a <- ggplot(data = nDEG, mapping = aes(x = Comparison, y = DEG, fill = Comparison)) + 
  geom_col() + 
  scale_fill_manual(values = c(colorRampPalette(colors = pal_tron()(2))(1), 
                               colorRampPalette(colors = pal_startrek()(4))(10)[c(3, 6, 9)])) + 
  theme_cowplot() + 
  theme(axis.title = element_text(size = 10),
        axis.text.x = element_blank(), 
        axis.ticks.x = element_blank(),
        legend.text = element_text(size = 8),
        legend.position = "top", 
        legend.justification = "center", 
        legend.direction = "horizontal", 
        legend.title = element_blank()) + 
  ylab(label = "DEGs") + 
  geom_text(aes(label = DEG)) + 
  guides(fill = guide_legend(override.aes = list(size = .5), ncol = 2))

# B) UMAP coloured by Clone
FigS3b <- ggplot(data = pheno, mapping = aes(x = UMAP_1, y = UMAP_2, color = Clone)) + 
  geom_point(size = .3) + 
  theme_cowplot() + 
  scale_color_tron() + 
  theme(axis.title = element_text(size = 10),
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        legend.text = element_text(size = 8),
        legend.position = "top", 
        legend.justification = "center", 
        legend.direction = "horizontal", 
        legend.title = element_blank()) +
  ggtitle(label = "") + 
  xlab(label = "UMAP 1") + 
  ylab(label = "UMAP 2") + 
  guides(colour = guide_legend(override.aes = list(size = 2)))

# C) Number of cells per Cluster coloured by Clone
FigS3c <- ggplot(data = pheno, mapping = aes(x = seurat_clusters, fill = Clone)) + 
  geom_bar() + 
  theme_cowplot() +
  theme(axis.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        legend.title = element_blank(),
        legend.position = "top", 
        legend.justification = "center", 
        legend.direction = "horizontal") + 
  ylab("Number of cells") + 
  xlab(label = "Cluster") + 
  scale_fill_tron() + 
  guides(fill = guide_legend(override.aes = list(size = .5)))

# D) Number of differentially expressed genes by Cluster
FigS3d <- ggplot(data = nDEG_Cluster, mapping = aes(x = Cluster, y = DEG, fill = Cluster)) + 
  geom_col() + 
  scale_fill_met_d(name = "Signac") + 
  theme_cowplot() +
  theme(axis.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        legend.title = element_blank(),
        legend.position = "top", 
        legend.justification = "center", 
        legend.direction = "horizontal") + 
  ylab("DEGs by Cluster") + 
  geom_text(aes(label = DEG)) +
  guides(fill = guide_legend(override.aes = list(size = .5), ncol = 4))

# E) Top genes per cluster 
FigS3e <- DotPlot(iDA, features = DEGs_Cluster_top, cluster.idents = TRUE) +
  theme(axis.title = element_text(size = 10), 
        axis.text.y = element_text(size = 6, hjust = 1, vjust = .5), 
        legend.title = element_blank(), 
        legend.position = "top", 
        legend.direction = "horizontal", 
        legend.box = "vertical") +
  coord_flip()
FigS3e$data$features.plot <- factor(gene_info$Gene_Name[match(x = DEGs_Cluster_top, table = gene_info$Gene_ID)], 
                                    levels = rev(gene_info$Gene_Name[match(x = DEGs_Cluster_top, table = gene_info$Gene_ID)]))

# F) Large list of known gene markers
FigS3f <- DotPlot(iDA, features = gm$Gene_ID, cluster.idents = TRUE) +
  theme(axis.title = element_text(size = 10), 
        axis.text.y = element_text(size = 6, hjust = 1, vjust = .5), 
        legend.title = element_blank(), 
        legend.position = "top", 
        legend.direction = "horizontal", 
        legend.box = "vertical") +
  coord_flip()
FigS3f$data$features.plot <- factor(gm$Gene_Name, levels = rev(gm$Gene_Name))

```

```{r FigSY}
# Comparison between D45 and D65.

# Dotplot 
#FigSY_a <- ggplot(data = DEA_Day65_Day45, mapping = aes(x = avg_log2FC, y = -log10(p_val_adj), label = Gene_Name)) + 
#  geom_point() + 
#  geom_text_repel() + theme_cowplot() +
#  theme(axis.text = element_text(size = 8),
#        legend.text = element_text(size = 8))


#FigSY_b <- DotPlot(iDA, features = rownames(DEA_Day65_Day45), cluster.idents = TRUE) 
#FigSY_b$data$features.plot <- factor(gene_info$Gene_Name[match(x = FigSY_b$data$features.plot,
#                                                               table = gene_info$Gene_ID)])
#FigSY_b <- FigSY_b + theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1, size = 8),
#                       legend.text = element_text(size = 8),
#                       legend.position = "top",
#                       legend.justification = "center",
#                       legend.direction = "horizontal")

## Comparison in just DA neurons.
#FigSY_DA_a <- ggplot(data = DEA_DA_Day65_Day45, mapping = aes(x = avg_log2FC, y = -log10(p_val_adj), label = Gene_Name)) + 
#  geom_point() + 
#  geom_text_repel() + theme_cowplot() +
#  theme(axis.text = element_text(size = 8),
#        legend.text = element_text(size = 8))


#FigSY_DA_b <- DotPlot(iDA_DA, features = rownames(DEA_DA_Day65_Day45), cluster.idents = FALSE) 
#FigSY_DA_b$data$features.plot <- factor(gene_info$Gene_Name[match(x = FigSY_DA_b$data$features.plot,
#                                                               table = gene_info$Gene_ID)])
#FigSY_DA_b <- FigSY_DA_b + theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1, size = 8),
#                       legend.text = element_text(size = 8),
#                       legend.position = "top",
#                       legend.justification = c(1, 4),
#                       legend.direction = "horizontal", 
#                       legend.box = "vertical") + coord_flip()

#FigSY_DA_c <- VlnPlot(iDA_DA, features = rownames(DEA_DA_Day65_Day45), stack = TRUE ,flip = TRUE, group.by = "Clone", split.by = "Day")
#FigSY_DA_c$data$feature <- factor(gene_info$Gene_Name[match(x = FigSY_DA_c$data$feature,
#                                                               table = gene_info$Gene_ID)])

```

```{r FigSX1}

# Optional plots Supplementary Figure X1.

# UMAP coloured by predicted cell type from the human midbrain
FigSXa <- ggplot(data = pheno, mapping = aes(x = UMAP_1, y = UMAP_2, color = Predicted)) +
  geom_point(size = .3) + 
  scale_color_uchicago() + 
  theme_cowplot() + 
  theme(axis.title = element_text(size = 10),
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        legend.text = element_text(size = 8),
        legend.position = "top", 
        legend.justification = "center", 
        legend.direction = "horizontal", 
        legend.title = element_blank()) +
  ggtitle(label = "") + 
  xlab(label = "UMAP 1") + 
  ylab(label = "UMAP 2") + 
  guides(colour = guide_legend(override.aes = list(size = 2)))

# UMAP showing the maximum prediction score for the human midbrain as reference
FigSXb <- ggplot(data = pheno, mapping = aes(x = UMAP_1, y = UMAP_2, color = Max_Prediction_Score)) +
  scale_color_material("deep-purple") + 
  geom_point(size = .3) + 
  theme_cowplot() + 
  theme(axis.title = element_text(size = 10),
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        legend.text = element_text(size = 8),
        legend.position = "top", 
        legend.justification = "center", 
        legend.direction = "horizontal", 
        legend.title = element_blank()) +
  ggtitle(label = "") + 
  xlab(label = "UMAP 1") + 
  ylab(label = "UMAP 2") 


```

```{r FigS3pdf}

#pdf(file = "./Figures/Fig3S.pdf", width = 7, height = 9)
#plot_grid(plot_grid(FigS3a, FigS3b, FigS3c, FigS3d, 
#                    rel_heights = c(1.1, 1, .9, 1.1), 
#                    ncol = 1, align = "hv", axis = "rtbl", 
#                    labels = c("A", "B", "C", "D")), 
#          FigS3e,
#          FigS3f, ncol = 3, labels = c("", "E", "F"))
#dev.off()

```

```{r Fig3pdf}

pdf(file = "./Figures/Fig3_draft.pdf", width = 7, height = 9)
plot_grid(plot_grid(Fig3a, Fig3b, Fig3c, ncol = 3, 
                    align = "h", axis = "bt", 
                    labels = c("A", "B", "C")),
          Fig3d,
          plot_grid(Fig3e, Fig3f, Fig3g, ncol = 3, 
                    align = "h", axis = "bt", 
                    labels = c("E", "F", "G")),
          ncol = 1, labels = c("", "D", ""), rel_heights = c(1, 1.1, 1))
dev.off()



```

```{r FigSX1pdf}

#pdf(file = "./Figures/FigSX.pdf", width = 7, height = 4)
#plot_grid(FigSXa, FigSXb, ncol = 2, labels = c("A", "B"))
#dev.off()


```

```{r OptionalFigs}

#FeaturePlot(object = iDA,
#            features = gene_info$Gene_ID[match(x = c("SYT1", "DCX", "TH"), 
#                                               table = gene_info$Gene_Name)])


# Dot plot some features #
#FeaturePlot(object = iDA, 
#            features = gene_info$Gene_ID[match(x = "FOXA2",table = gene_info$Gene_Name)])


#be <- c("PBX1")

#DAm <- c("TH", "DDC", "SLC6A3", "SLC18A2", "DRD2", "LMX1A", "LMX1B", "FOXA2", "NR4A2", "PITX3", "EN1", "EN2", "CNPY1", "PBX1", "GAD1", "GAD2")

#DAm <- c("TH", "NR4A2", "LMX1A", "LMX1B", "DDC", "KCNJ6", "PBX1", "LMO3", "CALB1", "SLC18A2", "TPH1", "SLC18A1", "SNAP25", 
#         "MAP2", "SYT1", "DCX", "VIM", "CORIN", "HES1", "NFIA", "RFX4", "SLC1A3", "SLIT2", "OTX2", "SOX6", "AGTR1", "ALDH1A1")

#prolif_markers <- c("MKI67", "TOP2A", "KIAA1524")
#prolif <- gene_info$Gene_ID[match(x = prolif_markers, table = gene_info$Gene_Name)]
#prolif <- prolif[prolif %in% rownames(iDA)]

#dp_prolif <- DotPlot(iDA, features = prolif, cluster.idents = TRUE) + coord_flip()
#dp_prolif$data$features.plot <- factor(gene_info$Gene_Name[match(x = dp_prolif$data$features.plot, table = gene_info$Gene_ID)], 
#                                levels = prolif_markers)


# Heatmap of top markers

# FigS3f <- DoHeatmap(object = iDA, features = DEGs_Cluster_top, group.by = "seurat_clusters", combine = FALSE)
# FigS3f <- FigS3f[[1]]
# FigS3f$data$Feature <- factor(gene_info$Gene_Name[match(FigS3f$data$Feature, gene_info$Gene_ID)],
# levels = rev(gene_info$Gene_Name[match(DEGs_Cluster_top, gene_info$Gene_ID)]))


# DEGs by Clone (obtained per cluster)

# DEGs_Clone <- unique(c(rownames(DEA_Clone_c0), rownames(DEA_Clone_c1), rownames(DEA_Clone_c2),
#                       rownames(DEA_Clone_c3), rownames(DEA_Clone_c4), rownames(DEA_Clone_c5),
#                       rownames(DEA_Clone_c6)))
# pfx <- DoHeatmap(object = iDA, features = DEGs_Clone, group.by = "Clone", group.colors = pal_tron()(2), angle = 0)
# pfx$data$Feature <- gene_info$Gene_Name[match(pfx$data$Feature, gene_info$Gene_ID)] 
# pfx 

# D) Number of differentially expressed genes by Clone per Cluster
#Fxd <- ggplot(data = nDEG_Clone, mapping = aes(x = Cluster, y = DEG, fill = Cluster)) + 
#  geom_col() + 
#  scale_fill_met_d(name = "Signac") + 
#  theme_cowplot() +
#  theme(legend.position = "top", 
#        legend.justification = "center", 
#        legend.direction = "horizontal", 
#        legend.title = element_blank(), 
#        legend.text = element_text(size = 8)) + 
#  ylab("DEGs by Clone") + 
#  geom_text(aes(label = DEG)) + 
#  guides(fill = guide_legend(override.aes = list(size = 4)))

```

```{r}

saveRDS(object = iDA, "./TimeCourse/TimeCourse_SeuratObject.rds")

saveRDS(object = DEA_Day30_Day21, "./TimeCourse/TimeCourse_DEA_Day30_Day21.rds")
saveRDS(object = DEA_Day45_Day30, "./TimeCourse/TimeCourse_DEA_Day45_Day30.rds")
saveRDS(object = DEA_Day65_Day45, "./TimeCourse/TimeCourse_DEA_Day65_Day45.rds")

saveRDS(object = DEA_Clone, "./TimeCourse/TimeCourse_DEA_Clone.rds")

saveRDS(object = DEA_c0, "./TimeCourse/TimeCourse_DEA_c0.rds")
saveRDS(object = DEA_c1, "./TimeCourse/TimeCourse_DEA_c1.rds")
saveRDS(object = DEA_c2, "./TimeCourse/TimeCourse_DEA_c0.rds")
saveRDS(object = DEA_c3, "./TimeCourse/TimeCourse_DEA_c0.rds")
saveRDS(object = DEA_c4, "./TimeCourse/TimeCourse_DEA_c0.rds")
saveRDS(object = DEA_c5, "./TimeCourse/TimeCourse_DEA_c0.rds")
saveRDS(object = DEA_c6, "./TimeCourse/TimeCourse_DEA_c0.rds")

saveRDS(object = DEGs_Cluster, "./TimeCourse/TimeCourse_DEGs_Cluster.rds")
saveRDS(object = iDA_pc_proj, "./TimeCourse/TimeCourse_iPSCprojection_hEM.rds")
saveRDS(object = hm_pcs, "./TimeCourse/TimeCourse_hEM_PCloadings.rds")


saveRDS(object = Fig3a, "./Figures/Fig3a.rds")
saveRDS(object = Fig3b, "./Figures/Fig3b.rds")
saveRDS(object = Fig3c, "./Figures/Fig3c.rds")
saveRDS(object = Fig3d, "./Figures/Fig3d.rds")
saveRDS(object = Fig3e, "./Figures/Fig3e.rds")
saveRDS(object = Fig3f, "./Figures/Fig3f.rds")
saveRDS(object = Fig3g, "./Figures/Fig3g.rds")


saveRDS(object = FigS3a, "./Figures/FigS3a.rds")
saveRDS(object = FigS3b, "./Figures/FigS3b.rds")
saveRDS(object = FigS3c, "./Figures/FigS3c.rds")
saveRDS(object = FigS3d, "./Figures/FigS3d.rds")
saveRDS(object = FigS3e, "./Figures/FigS3e.rds")
saveRDS(object = FigS3f, "./Figures/FigS3f.rds")

saveRDS(object = FigSXa, "./Figures/FigSXa.rds")
saveRDS(object = FigSXb, "./Figures/FigSXa.rds")

```


