---
title: "Filtering dataset and DEA iPSC-DA BFP sorting and MPP treatment"
author: "J Monzón Sandoval"
date: 'Last update: `r date()`'
output: 
  html_document:
    theme: lumen
    toc: yes
    number_sections: true
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

library(Seurat)
library(scater)
#library(ouija)

library(stringr)
library(ggplot2)
library(ggsci)
library(ggvenn)
library(ggnewscale)
library(cowplot)
library(MetBrewer)
library(RColorBrewer)

library(corrgram)
library(venn)
library(reshape)
library(pheatmap)
library(ComplexHeatmap)

library(clusterProfiler)

```

```{r GeneExpression}

# Gene_ID	Gene_Name	Gene_Biotype	Gene_Length # Columns in gene info tab 
# ensembl annotations release 99
gene_info <- read.table("./05_Analyzer/NS_basal_S15_L1/gene_info.csv", sep = ",", header = TRUE, stringsAsFactors = FALSE)

# All data matrices
fn <- list.files(path = "./05_Analyzer", 
                 pattern = "_genematrix.csv", 
                 recursive = TRUE, full.names = TRUE)
fn <- fn[str_detect(string = fn, pattern = "_incl_introns_", 
                    negate = TRUE)] # change to FALSE to include introns

# Need to combine counts from the different samples
t1 <- read.table(file = fn[1], sep = ",", header = TRUE, row.names = 1)

for (i in c(2:length(fn))){
  ti <- read.table(file = fn[i], sep = ",", header = TRUE, row.names = 1)
  t1 <- cbind(t1, ti)
}

ged <- t1
rm(i, t1, ti, fn)

ged <- as.matrix(ged)

```

```{r phenoPart1}

# Mappa Stats
pheno <- read.table("./06_Pheno/Pheno_MappaStats.txt", sep = "\t", stringsAsFactors = FALSE, header = TRUE)

pheno <- data.frame(pheno[, 1:12], 
                    Percent_Mapped = pheno$Mapped_Reads * 100 / pheno$Barcoded_Reads, 
                    Percent_Multimapped_Reads = pheno$Multimapped_Reads * 100 / pheno$Mapped_Reads,
                    Percent_Uniquely_Mapped = pheno$Uniquely_Mapped_Reads * 100 / pheno$Mapped_Reads,
                    Percent_Exon_Reads = pheno$Exon_Reads * 100 / pheno$Uniquely_Mapped,
                    Percent_Ambiguous_Exon_Reads = pheno$Ambiguous_Exon_Reads * 100 / pheno$Uniquely_Mapped,
                    Percent_Intron_Reads = pheno$Intron_Reads * 100 / pheno$Uniquely_Mapped,
                    Percent_Ambiguous_Intron_Reads = pheno$Ambiguous_Intron_Reads * 100 / pheno$Uniquely_Mapped,
                    Percent_Intergenic_Reads = pheno$Intergenic_Reads * 100 / pheno$Uniquely_Mapped,
                    Percent_Mitochondrial_Reads = pheno$Mitochondrial_Reads * 100 / pheno$Uniquely_Mapped,
                    Percent_Ribosomal_Reads = pheno$Ribosomal_Reads * 100 / pheno$Uniquely_Mapped,
                    pheno[, 13:ncol(pheno)])

# Set up levels 
pheno$InternalControl <- factor(x = pheno$InternalControl, levels = c("PosCtrl", "NegCtrl", "no"))
pheno$Status <- factor(x = pheno$Status, levels = c("Sorted", "NS", "PosCtrl", "NegCtrl"))
pheno$BFP <- factor(x = pheno$BFP, levels = c("NS", "negative", "positive",  "PosCtrl", "NegCtrl"))
pheno$Treatment <- factor(x = pheno$Treatment, levels = c("basal", "MPP", "PosCtrl", "NegCtrl"))
pheno$Dose <- factor(x = pheno$Dose, levels = c("0mM", "1mM", "PosCtrl", "NegCtrl"))
pheno$Time <- factor(x = pheno$Time, levels = c("0h", "1h", "6h", "12h", "24h", "PosCtrl", "NegCtrl"))
pheno$Group <- factor(x = pheno$Group, levels =  c("BFPplus_basal_0h", "BFPplus_MPP_1h", "BFPplus_MPP_6h", 
                                                   "BFPplus_MPP_12h", "BFPplus_MPP_24h", "BFPminus_basal_0h", 
                                                   "BFPminus_MPP_24h", "NS_basal_0h", "NS_MPP_24h", 
                                                   "PosCtrl", "NegCtrl"))
pheno$SampleDescription <- factor(x = pheno$SampleDescription, 
                                  levels =  c("Sorted BFP+ basal", "Sorted BFP+ 1mM MPP+ 1h", "Sorted BFP+ 1mM MPP+ 6h",
                                              "Sorted BFP+ 1mM MPP+ 12h", "Sorted BFP+ 1mM MPP+ 24h", "Sorted BFP- basal",
                                              "Sorted BFP- 1mM MPP+ 24h", "NS basal", "NS 1mM MPP+ 24h",
                                              "Positive control", "Negative control"))



```

```{r PhenoPart2}

# Feature annotation and counts only for Protein Coding genes
fd <- gene_info[which(gene_info$Gene_Biotype == "protein_coding"), ]
rownames(fd) <- fd$Gene_ID
pc_genes <- rownames(fd)

# Select protein coding genes and order of gene expression data to pheno
pc_counts <- ged[pc_genes, pheno$Barcode]

# Add row names to pheno to create sce object
pheno_sce <- pheno
rownames(pheno_sce) <- pheno_sce$Barcode

# single cell object
sce <- SingleCellExperiment(assays = list(counts = pc_counts),
                            colData = pheno_sce, rowData = fd)
rm(pheno_sce)


# Calculate QC for cells
QCcells <- perCellQCMetrics(sce, percent_top = c(10, 20, 50, 100)) 

# Normalize data 
cpm(sce) <- calculateCPM(sce)
tpm(sce) <- calculateTPM(sce, lengths = rowData(sce)[, "Gene_Length"])

# Add QC for cells and number of genes detected to colData
genes_detected_1TPM <- apply(X = tpm(sce), MARGIN = 2, FUN = function(x) length(which(x >= 1)))
genes_detected_5TPM <- apply(X = tpm(sce), MARGIN = 2, FUN = function(x) length(which(x >= 5)))
genes_detected_1cnt <- apply(X = counts(sce), MARGIN = 2, FUN = function(x) length(which(x >= 1)))
genes_detected_5cnt <- apply(X = counts(sce), MARGIN = 2, FUN = function(x) length(which(x >= 5)))

colData(sce) <- cbind(colData(sce), QCcells)
colData(sce) <- cbind(colData(sce), data.frame(genes_detected_1TPM = genes_detected_1TPM,
                                               genes_detected_5TPM = genes_detected_5TPM, 
                                               genes_detected_1cnt = genes_detected_1cnt,
                                               genes_detected_5cnt = genes_detected_5cnt))


rm(fd, QCcells)
rm(genes_detected_1TPM, genes_detected_5TPM, genes_detected_1cnt, genes_detected_5cnt)

# Cell pheno data
pheno <- as.data.frame(colData(sce), row.names = sce$Barcode)

rm(sce)

```

+ Total features: `r nrow(ged)`.

+ Total number of barcodes `r ncol(ged)`.

# Data filtering

## Protein coding genes

```{r}

# Seurat object
iDA <- CreateSeuratObject(counts = pc_counts, meta.data = pheno, project = "Sorted_Treatment", min.cells = 0, min.features = 0)

rm(ged, pc_counts, pc_genes)

```

+ Total features: `r nrow(iDA)` .

+ Total number of barcodes `r ncol(iDA)`.

## Filtering cells 

### Cells with low number of genes detected.

```{r fig.height=3, fig.width=4}

mingenes <- 2000

plot(density(pheno$genes_detected_1cnt), main = "Genes detected (1 count)")
abline( v = mingenes, col = "red", lty = 2)
text(x = mingenes, y = 0.00010, labels = paste(mingenes, "genes"), srt = 90, pos = 2)

```

```{r}

cat("\n + Barcodes per sample: \n")
table(pheno$Group)
cat("\n")

cat("\n + Barcodes per sample with < ", mingenes, "genes detected: \n")
table(pheno$Group[which(pheno$genes_detected_1cnt < mingenes)])
cat("\n")

cat("\n + Barcodes per sample with > ", mingenes, "genes detected: \n")
table(pheno$Group[which(pheno$genes_detected_1cnt > mingenes)])
cat("\n")

iDA <- subset(x = iDA, cells = pheno$Barcode[which(pheno$genes_detected_1cnt > mingenes)])

rm(mingenes)

```

+ Total features: `r nrow(iDA)` .

+ Total number of barcodes `r ncol(iDA)`.

### Internal controls

```{r}

cat("+ We furhter removed", length(which(iDA$Group == "NegCtrl")), " NegCtrl samples \n")
cat("+ We furhter removed", length(which(iDA$Group == "PosCtrl")), " PosCtrl samples \n")

iDA <- subset(x = iDA, cells = iDA$Barcode[!(iDA$Group == "NegCtrl")])
iDA <- subset(x = iDA, cells = iDA$Barcode[!(iDA$Group == "PosCtrl")])

```

+ Total features: `r nrow(iDA)` .

+ Total number of barcodes `r ncol(iDA)`.


### Cells with high % of mitochondrial reads

```{r}

cat("+ We furhter removed", length(which(iDA$Percent_Mitochondrial_Reads > 20)), 
    "cell with high % of mitochondrial reads (> 20%) \n")

iDA <- subset(x = iDA, cells = iDA$Barcode[!(iDA$Percent_Mitochondrial_Reads > 20)])

cat("At the end", sum(table(iDA$Group)), "cells remained:\t")
table(iDA$Group)

```

+ Total features: `r nrow(iDA)` .

+ Total number of barcodes `r ncol(iDA)`.

### Cells with low number of mapped reads

```{r}

cat("+ We furhter removed", length(which(iDA$Mapped_Reads < 100000)), 
    "cell with low numer of mapped reads (< 100000) \n")

iDA <- subset(x = iDA, cells = iDA$Barcode[!(iDA$Mapped_Reads < 100000)])

cat("At the end", sum(table(iDA$Group)), "cells remained:\t")
table(iDA$Group)

```

+ Total features: `r nrow(iDA)` .

+ Total number of barcodes `r ncol(iDA)`.

## Filtering genes

## Genes consistently expressed across cells

```{r warning=FALSE}

cells_nonzero <- apply(X = iDA[["RNA"]]@counts, MARGIN = 1, FUN = function(x) length(which(x != 0)))
cells_above1 <- apply(X = iDA[["RNA"]]@counts, MARGIN = 1, FUN = function(x) length(which(x > 1)))
cells_above5 <- apply(X = iDA[["RNA"]]@counts, MARGIN = 1, FUN = function(x) length(which(x > 5)))
cells_above10 <- apply(X = iDA[["RNA"]]@counts, MARGIN = 1, FUN = function(x) length(which(x > 10)))

output <- matrix(NA, nrow = 4, ncol = 100, dimnames = list(c("Non zero counts", "More than 1 count", "More than 5 counts", "More than 10 counts"), seq(.01, 1, .01)))

percent_cells <- seq(.01, 1, .01)

for(i in c(1:100)){
  output[1, i] <- length(which(cells_nonzero > ncol(iDA) * percent_cells[i]))
  output[2, i] <- length(which(cells_above1 > ncol(iDA) * percent_cells[i]))
  output[3, i] <- length(which(cells_above5 > ncol(iDA) * percent_cells[i]))
  output[4, i] <- length(which(cells_above10 > ncol(iDA) * percent_cells[i]))
  
}

rm(i, percent_cells)

output <- melt(output)
colnames(output) <- c("Counts", "Fractions_Of_cells", "Genes")

```


```{r fig.width=5.5, fig.height=3.5}

g1 <- ggplot(data = output, mapping = aes(x = Fractions_Of_cells, y = Genes, color = Counts)) + 
  geom_point(size = 1) + 
  scale_color_lancet() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) 

g1 

rm(g1, output)

```

## Filter lowly expressed genes

+ Keep genes with more than 1 count in at least 5 percent of the filtered cells.

```{r}

#keep_genes <- rownames(iDA)[which(cells_nonzero >= (ncol(iDA) * .1))]
keep_genes <- rownames(iDA)[which(cells_above1 >= (ncol(iDA) * .05))]
iDA <- subset(x = iDA, features = keep_genes)

rm(cells_above1, cells_above10, cells_above5, cells_nonzero)
rm(keep_genes)

# Drop levels from metadata
iDA@meta.data <- droplevels(iDA@meta.data)

```

+ Total features: `r nrow(iDA)` .

+ Total number of barcodes `r ncol(iDA)`.

## Gene markers

```{r genemarkers}

all_genes <- rownames(iDA)

# Cell type markers
gm <- read.table("../iPSC_DA_nuclei_cell_iCELL8/06_Pheno/CellType_Markers_Subset2.txt", sep = "\t", header = TRUE)
gm <- data.frame(gm, Gene_ID = gene_info$Gene_ID[match(gm$Gene_Name, gene_info$Gene_Name)])

cat("Markers not found in filtered data:", gm$Gene_Name[!(gm$Gene_ID %in% all_genes)],"\n")

gm <- gm[gm$Gene_ID %in% all_genes, ]

if(length(which(duplicated(gm$Gene_ID) == TRUE) > 0)) {
  gm <- gm[-which(duplicated(gm$Gene_ID) == TRUE), ]
}

gm$Gene_ID <- factor(gm$Gene_ID, levels = gm$Gene_ID)

```

# Basal

## Normalization and clustering 

+ Carried out in filtered cells.

+ Normalization LogNormalize (scale factor 1e6).

+ PCA based on the top 5000 most variable features.

+ UMAP based on top 10 PCS.

```{r fig.height=3, fig.width=4}

# Normalize subset of basal cells from Plate A
iDA_basal <- subset(x = iDA, cells = iDA$Barcode[(iDA$Plate == "A")])
iDA_basal <- subset(x = iDA_basal, cells = iDA_basal$Barcode[(iDA_basal$Treatment == "basal")])
iDA_basal <- NormalizeData(iDA_basal, normalization.method = "LogNormalize", scale.factor = 1e6)
iDA_basal <- FindVariableFeatures(iDA_basal, selection.method = "vst", nfeatures = 5000)

# Scale on all data run PC in top 5000 most variable features
iDA_basal <- ScaleData(iDA_basal, features = all_genes)
iDA_basal <- RunPCA(iDA_basal, features = VariableFeatures(object = iDA_basal), verbose = FALSE)
# Elbow plot
barplot(iDA_basal@reductions$pca@stdev[1:20], xlab = "PC", ylab = "Standard Deviation", 
        names.arg = 1:20, las = 2, col = grey.colors(20))

# Cell cycle score
s_genes <- gene_info$Gene_ID[match(cc.genes$s.genes, table = gene_info$Gene_Name)]
g2m_genes <- gene_info$Gene_ID[match(cc.genes$g2m.genes, table = gene_info$Gene_Name)]
s_genes <- intersect(rownames(iDA_basal), s_genes)
g2m_genes <- intersect(rownames(iDA_basal), g2m_genes)
iDA_basal <- CellCycleScoring(iDA_basal, s.features = s_genes, g2m.features = g2m_genes, set.ident = FALSE)
rm(g2m_genes, s_genes)

# UMAP and cell clustering 
iDA_basal <- RunUMAP(iDA_basal, reduction = "pca", dims = 1:10)
iDA_basal <- FindNeighbors(iDA_basal, dims = 1:10)
iDA_basal <- FindClusters(iDA_basal, resolution = .9)

# Rename clusters
iDA_basal$seurat_clusters <- as.factor(paste("B", iDA_basal$seurat_clusters, sep = ""))

# Threshold for DEGs
cut_adjp <- 0.05

```

## Cluster gene markers

+ Gene markers per cluster (log2FC > 0.25, adjusted p value < 0.05 and only positive true).

```{r DEAClusterbasal}

# Differential expression analysis: cluster markers # 

Idents(object = iDA_basal) <- iDA_basal$seurat_clusters

DEA_basal_b0 <- FindMarkers(object = iDA_basal, ident.1 = "B0",
                      logfc.threshold = 0.25, test.use = "wilcox", 
                      only.pos = TRUE)

DEA_basal_b1 <- FindMarkers(object = iDA_basal, ident.1 = "B1",
                      logfc.threshold = 0.25, test.use = "wilcox", 
                      only.pos = TRUE)

DEA_basal_b2 <- FindMarkers(object = iDA_basal, ident.1 = "B2",
                      logfc.threshold = 0.25, test.use = "wilcox", 
                      only.pos = TRUE)

DEA_basal_b3 <- FindMarkers(object = iDA_basal, ident.1 = "B3",
                      logfc.threshold = 0.25, test.use = "wilcox", 
                      only.pos = TRUE)

DEA_basal_b4 <- FindMarkers(object = iDA_basal, ident.1 = "B4",
                      logfc.threshold = 0.25, test.use = "wilcox", 
                      only.pos = TRUE)

DEA_basal_b5 <- FindMarkers(object = iDA_basal, ident.1 = "B5",
                      logfc.threshold = 0.25, test.use = "wilcox", 
                      only.pos = TRUE)

DEA_basal_b0 <- DEA_basal_b0[which(DEA_basal_b0$p_val_adj < cut_adjp), ]
DEA_basal_b1 <- DEA_basal_b1[which(DEA_basal_b1$p_val_adj < cut_adjp), ]
DEA_basal_b2 <- DEA_basal_b2[which(DEA_basal_b2$p_val_adj < cut_adjp), ]
DEA_basal_b3 <- DEA_basal_b3[which(DEA_basal_b3$p_val_adj < cut_adjp), ]
DEA_basal_b4 <- DEA_basal_b4[which(DEA_basal_b4$p_val_adj < cut_adjp), ]
DEA_basal_b5 <- DEA_basal_b5[which(DEA_basal_b5$p_val_adj < cut_adjp), ]

# Number of DEGs per Cluster
nDEG_Cluster_basal <- c(nrow(DEA_basal_b0), nrow(DEA_basal_b1), nrow(DEA_basal_b2),
                        nrow(DEA_basal_b3), nrow(DEA_basal_b4), nrow(DEA_basal_b5))

nDEG_Cluster_basal <- data.frame(Cluster = levels(iDA_basal$seurat_clusters),
                                 DEG = nDEG_Cluster_basal)

# Top DEGs by Cluster
DEGs_Cluster_top_basal <- unique(c(head(rownames(DEA_basal_b0), 10), head(rownames(DEA_basal_b1), 10),
                                   head(rownames(DEA_basal_b2), 10), head(rownames(DEA_basal_b3), 10),
                                   head(rownames(DEA_basal_b4), 10), head(rownames(DEA_basal_b5), 10))) #,

# All DEGs by Cluster
DEGs_Cluster_basal <- unique(c(rownames(DEA_basal_b0), rownames(DEA_basal_b1),
                               rownames(DEA_basal_b2), rownames(DEA_basal_b3),
                               rownames(DEA_basal_b4), rownames(DEA_basal_b5)))

# Add gene names
DEA_basal_b0 <- data.frame(DEA_basal_b0, 
                           Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_basal_b0),
                                                                 table = gene_info$Gene_ID)])

DEA_basal_b1 <- data.frame(DEA_basal_b1, 
                           Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_basal_b1),
                                                                 table = gene_info$Gene_ID)])

DEA_basal_b2 <- data.frame(DEA_basal_b2, 
                           Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_basal_b2),
                                                                 table = gene_info$Gene_ID)])

DEA_basal_b3 <- data.frame(DEA_basal_b3, 
                           Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_basal_b3),
                                                                 table = gene_info$Gene_ID)])

DEA_basal_b4 <- data.frame(DEA_basal_b4, 
                           Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_basal_b4),
                                                                 table = gene_info$Gene_ID)])

DEA_basal_b5 <- data.frame(DEA_basal_b5, 
                           Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_basal_b5),
                                                                 table = gene_info$Gene_ID)])

```

## Overlap cluster markers with cell type markers

```{r celltypeannotbasal, results='asis'}

# Overlap of cluster markers with cell type markers
cat("\n\n\ + Overlap with cluster 0 markers: \n")
gm[gm$Gene_ID %in% rownames(DEA_basal_b0),]
cat("\n\n\ + Overlap with cluster 1 markers: \n")
gm[gm$Gene_ID %in% rownames(DEA_basal_b1),]
cat("\n\n\ + Overlap with cluster 2 markers: \n")
gm[gm$Gene_ID %in% rownames(DEA_basal_b2),]
cat("\n\n\ + Overlap with cluster 3 markers: \n")
gm[gm$Gene_ID %in% rownames(DEA_basal_b3),]
cat("\n\n\ + Overlap with cluster 4 markers: \n")
gm[gm$Gene_ID %in% rownames(DEA_basal_b4),]
cat("\n\n\ + Overlap with cluster 5 markers: \n")
gm[gm$Gene_ID %in% rownames(DEA_basal_b5),]

# Overlap cluster markers and gene markers
ov_gm_DEGclust_basal <- gm[gm$Gene_ID %in% unique(unlist(DEGs_Cluster_basal)), ]
ov_gm_DEGclust_basal <- droplevels(ov_gm_DEGclust_basal)

```


## Cell type assignment

```{r celltypebasal}

# Assign cell clusters to cell types
cell_type <- as.character(iDA_basal$seurat_clusters)
cell_type[which(cell_type == "B0")] <- "Neuronal"
cell_type[which(cell_type == "B1")] <- "Neuronal"
cell_type[which(cell_type == "B2")] <- "Progenitor"
cell_type[which(cell_type == "B3")] <- "Neuronal"
cell_type[which(cell_type == "B4")] <- "Progenitor"
cell_type[which(cell_type == "B5")] <- "Glia"

cell_type <- factor(x = cell_type, 
                    levels = c("Progenitor", "Glia", "Neuronal"))

iDA_basal <- AddMetaData(object = iDA_basal, metadata = cell_type, col.name = "Cell_type")
rm(cell_type)

```


## Differential expression by Sorting status (BFP).

+ Contrasting gene expression only in neuronal clusters 

```{r}

iDA_basal_neu <- subset(x = iDA_basal, 
                        cells = iDA_basal$Barcode[which((iDA_basal$seurat_clusters %in% c("B0", "B1", "B3") == TRUE))])

Idents(object = iDA_basal_neu) <- iDA_basal_neu$BFP

DEA_pos_neg <- FindMarkers(object = iDA_basal_neu, ident.1 = "positive", ident.2 = "negative",
                           logfc.threshold = 0.25, test.use = "wilcox")

DEA_pos_NS <- FindMarkers(object = iDA_basal_neu, ident.1 = "positive", ident.2 = "NS",
                           logfc.threshold = 0.25, test.use = "wilcox")

DEA_neg_NS <- FindMarkers(object = iDA_basal_neu, ident.1 = "negative", ident.2 = "NS",
                           logfc.threshold = 0.25, test.use = "wilcox")

DEA_pos_neg <- DEA_pos_neg[which(DEA_pos_neg$p_val_adj < cut_adjp), ]
DEA_pos_NS <- DEA_pos_NS[which(DEA_pos_NS$p_val_adj < cut_adjp), ]
DEA_neg_NS <- DEA_neg_NS[which(DEA_neg_NS$p_val_adj < cut_adjp), ]
  
# Number of DEGs by Sorting status in the most neuronal fraction
nDEG_Sorting <- c(nrow(DEA_pos_neg), nrow(DEA_pos_NS), nrow(DEA_neg_NS))
nDEG_Sorting <- data.frame(Comparison = c("Positive_vs_Negative", 
                                          "Positive_vs_NS", 
                                          "Negative_vs_NS"), 
                           DEG = nDEG_Sorting)


DEGs_Sorting <- list(`BFP+ vs BFP-` = rownames(DEA_pos_neg),
                     `BFP+ vs NS` = rownames(DEA_pos_NS),
                     `BFP- vs NS` = rownames(DEA_neg_NS))

DEGs_Sorting_Type <- list(`UP_PositiveVsNegative_Neuronal` = rownames(DEA_pos_neg)[which(DEA_pos_neg$avg_log2FC > 0)],
                          `UP_PositiveVsNS_Neuronal` = rownames(DEA_pos_NS)[which(DEA_pos_NS$avg_log2FC > 0)],
                          `UP_NegativeVsNS_Neuronal` = rownames(DEA_neg_NS)[which(DEA_neg_NS$avg_log2FC > 0)],
                          `DOWN_PositiveVsNegative_Neuronal` = rownames(DEA_pos_neg)[which(DEA_pos_neg$avg_log2FC < 0)],
                          `DOWN_PositiveVsNS_Neuronal` = rownames(DEA_pos_NS)[which(DEA_pos_NS$avg_log2FC < 0)],
                          `DOWN_NegativeVsNS_Neuronal` = rownames(DEA_neg_NS)[which(DEA_neg_NS$avg_log2FC < 0)])


# Top 10 DEGS across the three comparisons.
DEGs_Sorting_Top10 <- unique(unlist(lapply(DEGs_Sorting, FUN = function(x) head(x, 10))))


# Add gene names
DEA_pos_neg <- data.frame(DEA_pos_neg, 
                          Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_pos_neg), 
                                                                table = gene_info$Gene_ID)])

DEA_pos_NS <- data.frame(DEA_pos_NS, 
                          Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_pos_NS), 
                                                                table = gene_info$Gene_ID)])

DEA_neg_NS <- data.frame(DEA_neg_NS, 
                          Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_neg_NS), 
                                                                table = gene_info$Gene_ID)])

```


```{r godegSort}

# Using cluster profiler

# Gene Ontology Analysis for two different p value thresholds
branch <- c("BP", "CC", "MF")
bg <- unique(gene_info$Gene_Name[match(rownames(iDA), table = gene_info$Gene_ID)])
cat("\n Background population:", length(bg), "\n")

for(j in c(1:length(branch))){
    for(i in c(1:length(DEGs_Sorting_Type))){
      genei <- DEGs_Sorting_Type[[i]]
      genei <- gene_info$Gene_Name[match(x = genei, table = gene_info$Gene_ID)]
      goi <- enrichGO(gene = genei, 
                      OrgDb = "org.Hs.eg.db",  
                      keyType = "SYMBOL",  
                      ont = branch[j],
                      pAdjustMethod = "fdr",
                      qvalueCutoff = 0.05,
                      universe = bg)
      saveRDS(goi, file = paste("./MPPSorted/GO_", 
                                           branch[j] ,"_DEAwilcox_", 
                                           names(DEGs_Sorting_Type)[i], ".rds", sep = ""))
      write.table(goi@result, file = paste("./MPPSorted/GO_", 
                                           branch[j] ,"_DEAwilcox_", 
                                           names(DEGs_Sorting_Type)[i], ".csv", sep = ""), 
                  sep = ",", quote = TRUE, row.names = FALSE)
      rm(goi, genei)
  }
}

rm(i, j, branch)

```


# MPP treated 

## Normalization and clustering 

+ Carried out in filtered cells.

+ Normalization LogNormalize (scale factor 1e6).

+ PCA based on the top 5000 most variable features.

+ UMAP based on top 10 PCS.

```{r fig.height=3, fig.width=4}

# Normalize subset of MPP cells from Plate A
iDA_MPP <- subset(x = iDA, cells = iDA$Barcode[(iDA$Plate == "A")])
iDA_MPP <- subset(x = iDA_MPP, cells = iDA_MPP$Barcode[(iDA_MPP$Treatment == "MPP")])
iDA_MPP <- NormalizeData(iDA_MPP, normalization.method = "LogNormalize", scale.factor = 1e6)
iDA_MPP <- FindVariableFeatures(iDA_MPP, selection.method = "vst", nfeatures = 5000)
all_genes <- rownames(iDA_MPP)

# Scale on all data run PC in top 5000 most variable features
iDA_MPP <- ScaleData(iDA_MPP, features = all_genes)
iDA_MPP <- RunPCA(iDA_MPP, features = VariableFeatures(object = iDA_MPP), verbose = FALSE)
# Elbow plot
barplot(iDA_MPP@reductions$pca@stdev[1:20], xlab = "PC", ylab = "Standard Deviation", 
        names.arg = 1:20, las = 2, col = grey.colors(20))

# Cell cycle score
s_genes <- gene_info$Gene_ID[match(cc.genes$s.genes, table = gene_info$Gene_Name)]
g2m_genes <- gene_info$Gene_ID[match(cc.genes$g2m.genes, table = gene_info$Gene_Name)]
s_genes <- intersect(rownames(iDA_MPP), s_genes)
g2m_genes <- intersect(rownames(iDA_MPP), g2m_genes)
iDA_MPP <- CellCycleScoring(iDA_MPP, s.features = s_genes, g2m.features = g2m_genes, set.ident = FALSE)
rm(g2m_genes, s_genes)

# UMAP and cell clustering 
iDA_MPP <- RunUMAP(iDA_MPP, reduction = "pca", dims = 1:10)
iDA_MPP <- FindNeighbors(iDA_MPP, dims = 1:10)
iDA_MPP <- FindClusters(iDA_MPP, resolution = 1.2)

# Rename clusters
iDA_MPP$seurat_clusters <- as.factor(paste("M", iDA_MPP$seurat_clusters, sep = ""))

# Threshold for DEGs
cut_adjp <- 0.05

```


## Cluster gene markers

+ Gene markers per cluster (log2FC > 0.25, adjusted p value < 0.05 and only positive true).

```{r DEAClusterMPP}

# Differential expression analysis: cluster markers # 

Idents(object = iDA_MPP) <- iDA_MPP$seurat_clusters

DEA_MPP_m0 <- FindMarkers(object = iDA_MPP, ident.1 = "M0",
                      logfc.threshold = 0.25, test.use = "wilcox", 
                      only.pos = TRUE)

DEA_MPP_m1 <- FindMarkers(object = iDA_MPP, ident.1 = "M1",
                      logfc.threshold = 0.25, test.use = "wilcox", 
                      only.pos = TRUE)

DEA_MPP_m2 <- FindMarkers(object = iDA_MPP, ident.1 = "M2",
                      logfc.threshold = 0.25, test.use = "wilcox", 
                      only.pos = TRUE)

DEA_MPP_m3 <- FindMarkers(object = iDA_MPP, ident.1 = "M3",
                      logfc.threshold = 0.25, test.use = "wilcox", 
                      only.pos = TRUE)

DEA_MPP_m4 <- FindMarkers(object = iDA_MPP, ident.1 = "M4",
                      logfc.threshold = 0.25, test.use = "wilcox", 
                      only.pos = TRUE)

DEA_MPP_m5 <- FindMarkers(object = iDA_MPP, ident.1 = "M5",
                      logfc.threshold = 0.25, test.use = "wilcox", 
                      only.pos = TRUE)

DEA_MPP_m0 <- DEA_MPP_m0[which(DEA_MPP_m0$p_val_adj < cut_adjp), ]
DEA_MPP_m1 <- DEA_MPP_m1[which(DEA_MPP_m1$p_val_adj < cut_adjp), ]
DEA_MPP_m2 <- DEA_MPP_m2[which(DEA_MPP_m2$p_val_adj < cut_adjp), ]
DEA_MPP_m3 <- DEA_MPP_m3[which(DEA_MPP_m3$p_val_adj < cut_adjp), ]
DEA_MPP_m4 <- DEA_MPP_m4[which(DEA_MPP_m4$p_val_adj < cut_adjp), ]
DEA_MPP_m5 <- DEA_MPP_m5[which(DEA_MPP_m5$p_val_adj < cut_adjp), ]

# Number of DEGs per Cluster
nDEG_Cluster_MPP <- c(nrow(DEA_MPP_m0), nrow(DEA_MPP_m1), nrow(DEA_MPP_m2),
                      nrow(DEA_MPP_m3), nrow(DEA_MPP_m4), nrow(DEA_MPP_m5))

nDEG_Cluster_MPP <- data.frame(Cluster = levels(iDA_MPP), #
                               DEG = nDEG_Cluster_MPP)

# Top DEGs by Cluster
DEGs_Cluster_top_MPP <- unique(c(head(rownames(DEA_MPP_m0), 10), head(rownames(DEA_MPP_m1), 10),
                                 head(rownames(DEA_MPP_m2), 10), head(rownames(DEA_MPP_m3), 10),
                                 head(rownames(DEA_MPP_m4), 10), head(rownames(DEA_MPP_m5), 10)))

# All DEGs by Cluster
DEGs_Cluster_MPP <- unique(c(rownames(DEA_MPP_m0), rownames(DEA_MPP_m1),
                             rownames(DEA_MPP_m2), rownames(DEA_MPP_m3),
                             rownames(DEA_MPP_m4), rownames(DEA_MPP_m5)))

# Add gene names
DEA_MPP_m0 <- data.frame(DEA_MPP_m0, 
                           Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_MPP_m0),
                                                                 table = gene_info$Gene_ID)])

DEA_MPP_m1 <- data.frame(DEA_MPP_m1, 
                           Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_MPP_m1),
                                                                 table = gene_info$Gene_ID)])

DEA_MPP_m2 <- data.frame(DEA_MPP_m2, 
                           Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_MPP_m2),
                                                                 table = gene_info$Gene_ID)])

DEA_MPP_m3 <- data.frame(DEA_MPP_m3, 
                           Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_MPP_m3),
                                                                 table = gene_info$Gene_ID)])

DEA_MPP_m4 <- data.frame(DEA_MPP_m4, 
                           Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_MPP_m4),
                                                                 table = gene_info$Gene_ID)])

DEA_MPP_m5 <- data.frame(DEA_MPP_m5, 
                           Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_MPP_m5),
                                                                 table = gene_info$Gene_ID)])

```

## Overlap cluster markers with cell type markers

```{r celltypeannot, results='asis'}

# Overlap of cluster markers with cell type markers
cat("\n\n\ + Overlap with cluster 0 markers: \n")
gm[gm$Gene_ID %in% rownames(DEA_MPP_m0),]
cat("\n\n\ + Overlap with cluster 1 markers: \n")
gm[gm$Gene_ID %in% rownames(DEA_MPP_m1),]
cat("\n\n\ + Overlap with cluster 2 markers: \n")
gm[gm$Gene_ID %in% rownames(DEA_MPP_m2),]
cat("\n\n\ + Overlap with cluster 3 markers: \n")
gm[gm$Gene_ID %in% rownames(DEA_MPP_m3),]
cat("\n\n\ + Overlap with cluster 4 markers: \n")
gm[gm$Gene_ID %in% rownames(DEA_MPP_m4),]
cat("\n\n\ + Overlap with cluster 5 markers: \n")
gm[gm$Gene_ID %in% rownames(DEA_MPP_m5),]

# Overlap cluster markers and gene markers
ov_gm_DEGclust_MPP <- gm[gm$Gene_ID %in% unique(unlist(DEGs_Cluster_MPP)), ]
ov_gm_DEGclust_MPP <- droplevels(ov_gm_DEGclust_MPP)


```

## Cell type assignment

```{r celltypeMPP}

# Assign cell clusters to cell types
cell_type <- as.character(iDA_MPP$seurat_clusters)
cell_type[which(cell_type == "M0")] <- "Neuronal"
cell_type[which(cell_type == "M1")] <- "Progenitor"
cell_type[which(cell_type == "M2")] <- "Neuronal"
cell_type[which(cell_type == "M3")] <- "Neuronal"
cell_type[which(cell_type == "M4")] <- "Progenitor"
cell_type[which(cell_type == "M5")] <- "Progenitor"

cell_type <- factor(x = cell_type, 
                    levels = c("Progenitor", "Neuronal"))

iDA_MPP <- AddMetaData(object = iDA_MPP, metadata = cell_type, col.name = "Cell_type")
rm(cell_type)

```

## Prediction using Basal clusters

```{r}

# Prediction of MPP clusters based on basal 
tranf_anchors <- FindTransferAnchors(reference = iDA_basal, 
                                     query = iDA_MPP, 
                                     dims = 1:10, 
                                     reference.reduction = "pca")

predictions <- TransferData(anchorset = tranf_anchors, 
                            refdata = iDA_basal$seurat_clusters, 
                            dims = 1:10)

iDA_MPP <- AddMetaData(object = iDA_MPP, 
                       metadata = predictions$predicted.id,
                       col.name = "Predicted_Basal")

iDA_MPP <- AddMetaData(object = iDA_MPP, 
                       metadata = predictions$prediction.score.max,
                       col.name = "Max_Prediction_Score")

rm(tranf_anchors, predictions)

```

## Proportion basal/predicted basal clusters before and after MPP

```{r}

# Proportion of basal and predicted (for MPP) basal cell clusters.
prop_clusters <- data.frame(Cluster = levels(iDA_basal$seurat_clusters),
                            Basal = as.numeric(table(iDA_basal$seurat_clusters)) / ncol(iDA_basal), 
                            MPP = as.numeric(table(iDA_MPP$Predicted_Basal)) / ncol(iDA_MPP))

# Number of cells per cluster (basal/predicted basal)
cell_cluster <- data.frame(rbind(as.numeric(table(iDA_basal$seurat_clusters)),
                                 as.numeric(table(iDA_MPP$Predicted_Basal))), 
                           row.names = c("Basal", "MPP"))
colnames(cell_cluster) <- levels(iDA_basal$seurat_clusters)


# Number of cells per neuronal cluster in Basal and MPP (predicted basal cluster)
cat("\n\n+ Cells per neuronal cluster (basal/ predicted basal):\n")
cell_cluster[, c("B0", "B1", "B3")]
cat("\n")

# Test if the proportions in the neuronal populations are different between basal and MPP
chisq.test(cell_cluster[, c("B0", "B1", "B3")]) 

# Basal cluster 0, 1 and 3
# X-squared = 60.517, df = 2, p-value = 7.224e-14
```

## Number of NS, BFP positive and BFP negative neurons before and after MPP

+ In neuronal clusters B0, B1, B3, M0, M2, M3.

```{r}

# Number of neuronal NS, BFP+ and BFP- cells.
cell_BFP <- data.frame(rbind(as.numeric(table(iDA_basal$BFP[which(iDA_basal$Cell_type %in% "Neuronal")])),
                             as.numeric(table(iDA_MPP$BFP[which(iDA_MPP$Cell_type %in% "Neuronal")]))),
                       row.names = c("Basal", "MPP"))
colnames(cell_BFP) <- levels(iDA_basal$BFP)


# Number of cells in neuronal clusters in Basal and MPP (predicted basal cluster) per sorting status
cat("\n\n+ Neuronal cells per sorting status in Basal and MPP+ treated\n")
cell_BFP
cat("\n")


# Test if the proportions of neuronal NS, BFP+ and BFP- populations are different between basal and MPP
chisq.test(cell_BFP)

#	Pearson's Chi-squared test
# data:  cell_BFP
# X-squared = 14.118, df = 2, p-value = 0.0008598

# Old data # X-squared = 6.6959, df = 2, p-value = 0.03516 # Old data #

```


# Basal and MPP

+ Carried out in filtered cells.

+ Normalization LogNormalize (scale factor 1e6).

+ PCA based on the top 5000 most variable features.

+ UMAP based on top 10 PCS.

```{r fig.height=3, fig.width=4}

# Normalize subset of basal cells from Plate A
iDA_both <- subset(x = iDA, cells = iDA$Barcode[(iDA$Plate == "A")])
iDA_both <- NormalizeData(iDA_both, normalization.method = "LogNormalize", scale.factor = 1e6)
iDA_both <- FindVariableFeatures(iDA_both, selection.method = "vst", nfeatures = 5000)
all_genes <- rownames(iDA_both)

# Scale on all data run PC in top 5000 most variable features
iDA_both <- ScaleData(iDA_both, features = all_genes)
iDA_both <- RunPCA(iDA_both, features = VariableFeatures(object = iDA_both), verbose = FALSE)

# Elbow plot
barplot(iDA_both@reductions$pca@stdev[1:20], xlab = "PC", ylab = "Standard Deviation", 
        names.arg = 1:20, las = 2, col = grey.colors(20))

# UMAP
iDA_both <- RunUMAP(iDA_both, reduction = "pca", dims = 1:10)

# Add cluster annotations 
cluster_identity <- c(iDA_basal$seurat_clusters, iDA_MPP$seurat_clusters)
cluster_identity <- cluster_identity[colnames(iDA_both)]

cell_type <- c(iDA_basal$Cell_type, iDA_MPP$Cell_type)
cell_type <- cell_type[colnames(iDA_both)]

# Add cluster annotation and cell type 
iDA_both <- AddMetaData(object = iDA_both, metadata = cluster_identity, col.name = "seurat_clusters")
iDA_both <- AddMetaData(object = iDA_both, metadata = cell_type, col.name = "Cell_type")

rm(cluster_identity, cell_type)

# Threshold for DEGs
cut_adjp <- 0.05

```

## Differential expression MPP vs basal

+ In neuronal cells by sorting status.

```{r}

# Neuronal cells only
iDA_both_neu <- subset(iDA_both, cells = iDA_both$Barcode[which(iDA_both$Cell_type %in% "Neuronal")])

# Create subgroup for contrast
subgroup <- paste(iDA_both_neu$BFP, iDA_both_neu$Treatment, sep = "_")
iDA_both_neu <- AddMetaData(object = iDA_both_neu, metadata = subgroup, col.name = "Subgroup")

Idents(object = iDA_both_neu) <- iDA_both_neu$Subgroup

DEA_MPP_basal_positive <- FindMarkers(object = iDA_both_neu, ident.1 = "positive_MPP", ident.2 = "positive_basal",
                                      logfc.threshold = 0.25, test.use = "wilcox")

DEA_MPP_basal_negative <- FindMarkers(object = iDA_both_neu, ident.1 = "negative_MPP", ident.2 = "negative_basal",
                                      logfc.threshold = 0.25, test.use = "wilcox")

DEA_MPP_basal_NS <- FindMarkers(object = iDA_both_neu, ident.1 = "NS_MPP", ident.2 = "NS_basal",
                                      logfc.threshold = 0.25, test.use = "wilcox")


DEA_MPP_basal_positive <- DEA_MPP_basal_positive[which(DEA_MPP_basal_positive$p_val_adj < cut_adjp), ]
DEA_MPP_basal_negative <- DEA_MPP_basal_negative[which(DEA_MPP_basal_negative$p_val_adj < cut_adjp), ]
DEA_MPP_basal_NS <- DEA_MPP_basal_NS[which(DEA_MPP_basal_NS$p_val_adj < cut_adjp), ]
  
# Number of DEGs between MPP and basal neurons (Done separately per sorting status)
nDEG_MPP <- c(nrow(DEA_MPP_basal_positive), nrow(DEA_MPP_basal_negative), nrow(DEA_MPP_basal_NS))
nDEG_MPP <- data.frame(Comparison = c("MPP_vs_basal_positive", 
                                      "MPP_vs_basal_negative", 
                                      "MPP_vs_basal_NS"), 
                           DEG = nDEG_MPP)

DEGs_MPP <- list(`MPPvsBasalPositive` = rownames(DEA_MPP_basal_positive),
                 `MPPvsBasalNegative` = rownames(DEA_MPP_basal_negative),
                 `MPPvsBasalNS` = rownames(DEA_MPP_basal_NS))

# Top 20 DEGS across the three comparisons.
DEGs_MPP_Top20 <- unique(unlist(lapply(DEGs_MPP, FUN = function(x) head(x, 20))))


# Add gene names
DEA_MPP_basal_positive <- data.frame(DEA_MPP_basal_positive,
                                     Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_MPP_basal_positive),
                                                                           table = gene_info$Gene_ID)])

DEA_MPP_basal_negative <- data.frame(DEA_MPP_basal_negative,
                                     Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_MPP_basal_negative),
                                                                           table = gene_info$Gene_ID)])

DEA_MPP_basal_NS <- data.frame(DEA_MPP_basal_NS,
                               Gene_Name = gene_info$Gene_Name[match(x = rownames(DEA_MPP_basal_NS),
                                                                     table = gene_info$Gene_ID)])


```


```{r godegMPP}

# Using cluster profiler

# Gene Ontology Analysis for two different p value thresholds
branch <- c("BP", "CC", "MF")
bg <- unique(gene_info$Gene_Name[match(rownames(iDA), table = gene_info$Gene_ID)])
cat("\n Background population:", length(bg), "\n")

for(j in c(1:length(branch))){
    for(i in c(1:length(DEGs_MPP))){
      genei <- DEGs_MPP[[i]]
      genei <- gene_info$Gene_Name[match(x = genei, table = gene_info$Gene_ID)]
      goi <- enrichGO(gene = genei, 
                      OrgDb = "org.Hs.eg.db",  
                      keyType = "SYMBOL",  
                      ont = branch[j],
                      pAdjustMethod = "fdr",
                      qvalueCutoff = 0.05,
                      universe = bg)
      saveRDS(goi, file = paste("./MPPSorted/GO_", 
                                branch[j] ,"_DEAwilcox_", 
                                names(DEGs_MPP)[i], ".rds", sep = ""))

      write.table(goi@result, file = paste("./MPPSorted/GO_", 
                                           branch[j] ,"_DEAwilcox_", 
                                           names(DEGs_MPP)[i], ".csv", sep = ""), 
                  sep = ",", quote = TRUE, row.names = FALSE)
      rm(goi, genei)
  }
}

rm(i, j, branch)

```


## VTA and SN DEG (Nichterwitz 2016)

```{r VTAandSN}

# 111 genes differentially expressed between the SN and the VTA
vta_sn <- read.table(file = "./07_GeneMarkers/VTA_SN_DEG_Nichterwitz_2016.txt", header = TRUE)
vta_sn <- vta_sn[vta_sn$Ensembl %in% rownames(iDA), ]
area <- vector()
area[which(vta_sn$log2FoldChange > 0)] <- "VTA"
area[which(vta_sn$log2FoldChange < 0)] <- "SN"
vta_sn <- data.frame(vta_sn, area = area, row.names = 1:nrow(vta_sn))
rm(area)

```

```{r VTASNheatmap}

#iDA_both_neu

mat <- iDA_both_neu[["RNA"]]@data[vta_sn$Ensembl, ] %>% as.matrix()

## scale the rows
mat <- t(scale(t(mat)))

## quantile(mat, c(0.1, 0.95))
cluster_an1 <- iDA_both_neu$Treatment
cluster_an2 <- iDA_both_neu$BFP
cluster_an3 <- droplevels(iDA_both_neu$seurat_clusters)

clustrows <- factor(paste(cluster_an1, cluster_an2, cluster_an3, sep = "_"), 
               levels = c("basal_NS_B0", "basal_NS_B1", "basal_NS_B3",  
                          "basal_positive_B0", "basal_positive_B1", "basal_positive_B3",
                          "basal_negative_B0",
                          "MPP_NS_M0", "MPP_NS_M2", 
                          "MPP_positive_M0", "MPP_positive_M2", "MPP_positive_M3",
                          "MPP_negative_M2", "MPP_negative_M3"))


col_fun = circlize::colorRamp2(c(-1.5, 0, 1.5), c("#FF00FF", "black", "#FFFF00"))


pdf(file = "./Figures/VTA_SN_markers_heatmap_Neurons.pdf", width = 7, height = 9)

Heatmap(mat, name = "Expression",  
        column_split = factor(clustrows),
        row_split = vta_sn$area,
        row_labels = vta_sn$RefSeq, 
        cluster_columns = FALSE,
        show_column_dend = FALSE,
        cluster_column_slices = TRUE,
        column_title_gp = gpar(fontsize = 8),
        column_gap = unit(0.5, "mm"),
        cluster_rows = TRUE,
        show_row_dend = FALSE,
        col = col_fun,
        row_names_gp = gpar(fontsize = 4),
        column_title_rot = 90,
        top_annotation = HeatmapAnnotation(Treatment = cluster_an1, 
                                           BFP = cluster_an2, 
                                           Cluster = cluster_an3,
                                           col = list(Treatment = c("basal" = pal_aaas()(2)[1], "MPP" = pal_aaas()(2)[2]),
                                                      BFP = c("NS" = unlist(as.list(met.brewer(name = "Isfahan2", n = 3)))[1],
                                                              "negative" = unlist(as.list(met.brewer(name = "Isfahan2", n = 3)))[2], 
                                                              "positive" = unlist(as.list(met.brewer(name = "Isfahan2", n = 3)))[3]),
                                                      Cluster = c("B0" = pal_d3()(6)[1], "B1" = pal_d3()(6)[2], "B3" = pal_d3()(6)[4], 
                                                                  "M0" = unlist(as.list(met.brewer(name = "VanGogh2", n = 6)))[1], 
                                                                  "M2" = unlist(as.list(met.brewer(name = "VanGogh2", n = 6)))[3],  
                                                                  "M3" = unlist(as.list(met.brewer(name = "VanGogh2", n = 6)))[4]))),
        show_column_names = FALSE,
        use_raster = TRUE,
        raster_quality = 4)

dev.off()

rm(mat, clustrows, col_fun, cluster_an1, cluster_an2, cluster_an3)

```

```{r ovVTASN}

cat("\n\n Overlap VTA vs SN and BFP Pos vs Neg degs: \n")
gene_info$Gene_Name[match(x = intersect(rownames(DEA_pos_neg), vta_sn$Ensembl), table = gene_info$Gene_ID)]

DEA_pos_neg[intersect(rownames(DEA_pos_neg), vta_sn$Ensembl),]
cat("\n")

ov <- intersect(rownames(DEA_pos_neg), vta_sn$Ensembl)
figov <- DotPlot(object = iDA_basal_neu, features = ov)
figov$data$features.plot <- gene_info$Gene_Name[match(x = figov$data$features.plot, table = gene_info$Gene_ID)]
figov <- figov  +theme(legend.position = "top", legend.direction = "horizontal", legend.box = "vertical", legend.title = element_blank())
figov2 <- VlnPlot(object = iDA_basal_neu, features = ov, same.y.lims = TRUE, combine = FALSE)
figov2 <- lapply(figov2, FUN = function(x) x + ggtitle(label = gene_info$Gene_Name[match(x$labels$title, table = gene_info$Gene_ID)]))

pdf("./Figures/VTA_SN_Neurons_Basal.pdf", height = 7, width = 7)
plot_grid(figov, figov2[[1]], figov2[[2]], figov2[[3]], ncol = 2)
dev.off()

rm(figov, figov2)

```


## VTA and SN (Del Aguila 2021)

```{r}

sn <- c("GSG1L", "ATP2A3", "CBLN1", "RGS16", "SLIT1")
vta <- c("CADM1", "NECAB1",  "TIMP2", "GNG4", "FXYD6", 
         "ZCCHC12", "KCNIP4", "CDH13", "OSBPL3", "ARHGAP26", 
         "PEG3", "CRYM", "SERPINE2", "PCSK2") # "LYH6" not matched # EN2 not in dataset

sn_vta_ens <- gene_info$Gene_ID[match(c(sn, vta), table = gene_info$Gene_Name)]

# Markers by Del Aguila
cat("\n\n + SN gene markers: \n")
sn
cat("\n\n + VTA gene markers: \n")
vta

# Overlap with genes DEG bewteen sorted populations
cat("\n\n + Ovelapping markers DE between Pos and Neg \n")
gene_info$Gene_Name[match(rownames(DEA_pos_neg)[rownames(DEA_pos_neg) %in% sn_vta_ens], gene_info$Gene_ID)]
cat("\n")

```


# Code for figures

```{r obj4plots}

# Basal 

# Updated pheno data for Figs # Add UMAP dims
pheno_basal <- data.frame(iDA_basal@meta.data, iDA_basal@reductions$umap@cell.embeddings[, 1:2])


# MPP 

# Updated pheno data for Figs # Add UMAP dims
pheno_MPP <- data.frame(iDA_MPP@meta.data, iDA_MPP@reductions$umap@cell.embeddings[, 1:2])


# Both

# Updated pheno data for Figs # Add UMAP dims
pheno_both <- data.frame(iDA_both@meta.data, iDA_both@reductions$umap@cell.embeddings[, 1:2])


# DEGs in MPP vs basal neurons.
DEGs_MPP <- list(`MPP vs Basal\nBFP+` = rownames(DEA_MPP_basal_positive),
                 `MPP vs Basal\nBFP-` = rownames(DEA_MPP_basal_negative),
                 `MPP vs Basal\nNS` = rownames(DEA_MPP_basal_NS))


```

```{r fig5}

# UMAP coloured by cluster
cls <- aggregate.data.frame(x = pheno_basal[c("UMAP_1", "UMAP_2")], 
                            by = list(pheno_basal$seurat_clusters), 
                            FUN = median)
cls <- data.frame(cls, Cell_type = pheno_basal$Cell_type[match(cls$Group.1, table = pheno_basal$seurat_clusters)])

Fig5f <- ggplot(data = pheno_basal, mapping = aes(x = UMAP_1, y = UMAP_2, color = seurat_clusters)) +
  geom_point(size = .3) + 
  scale_color_d3(alpha = .7) + 
  theme_cowplot() + 
  theme(axis.title = element_text(size = 10),
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        legend.key.size = unit(x = .3, units = "cm"),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
        legend.spacing.x = unit(.5, "mm"),
        legend.spacing.y = unit(2, "mm"),
        legend.text = element_text(size = 8),
        legend.position = "top", 
        legend.justification = "center", 
        legend.direction = "horizontal", 
        legend.title = element_blank()) +
  ggtitle(label = "") + 
  xlab(label = "UMAP 1") + 
  ylab(label = "UMAP 2") + 
  guides(colour = guide_legend(override.aes = list(size = 2))) 
Fig5f <- Fig5f + annotate("text", x = cls$UMAP_1, y = cls$UMAP_2, label = cls$Cell_type, size = 2)


# A few markers
mkr <- c("HMGB1", "HMGB2", "NR2F1", "ZEB2", "CDH2", "SOX9", "SNAP25", "SYT1", "NEFL", "NEFM", "SST", "MAOA", "OTX2")
Fig5g <- DotPlot(object = iDA_basal, 
                 features = gene_info$Gene_ID[match(x = mkr, table = gene_info$Gene_Name)], 
                 dot.scale = 3) +
  theme(axis.title = element_text(size = 10), 
        axis.text.y = element_text(size = 8, hjust = 1, vjust = .5),
        axis.text.x = element_text(size = 8),
        legend.key.size = unit(x = .3, units = "cm"),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
        legend.spacing.x = unit(.5, "mm"),
        legend.spacing.y = unit(2, "mm"),
        legend.text = element_text(size = 8),
        legend.title = element_blank(), 
        legend.position = "top", 
        legend.direction = "horizontal",
        legend.box = "vertical") +
  coord_flip()

Fig5g$data$features.plot <- factor(gene_info$Gene_Name[match(x = Fig5g$data$features.plot, table = gene_info$Gene_ID)], 
                                   levels = rev(mkr))

# UMAP coloured by sorting status
Fig5h <- ggplot(data = pheno_basal, mapping = aes(x = UMAP_1, y = UMAP_2, color = BFP)) +
  geom_point(size = .3) + 
  scale_color_met_d(name = "Isfahan2") + 
  theme_cowplot() + 
  theme(axis.title = element_text(size = 10),
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        legend.key.size = unit(x = .3, units = "cm"),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
        legend.spacing.x = unit(.5, "mm"),
        legend.spacing.y = unit(2, "mm"),
        legend.text = element_text(size = 8),
        legend.position = "top", 
        legend.justification = "center", 
        legend.direction = "horizontal", 
        legend.title = element_blank()) +
  ggtitle(label = "") + 
  xlab(label = "UMAP 1") + 
  ylab(label = "UMAP 2") + 
  guides(colour = guide_legend(override.aes = list(size = 2))) 


# Number of cells per Cluster coloured sorting status
Fig5i <- ggplot(data = pheno_basal, mapping = aes(x = seurat_clusters, fill = BFP)) + 
  geom_bar() + 
  theme_cowplot() +
  theme(axis.title = element_text(size = 10),
        axis.text = element_text(size = 8),
        legend.key.size = unit(x = .3, units = "cm"),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
        legend.spacing.x = unit(.5, "mm"),
        legend.spacing.y = unit(2, "mm"),
        legend.text = element_text(size = 8),
        legend.title = element_blank(),
        legend.position = "top", 
        legend.justification = "center", 
        legend.direction = "vertical") + 
  ylab("Number of cells") + 
  xlab(label = "Cluster") + 
  scale_fill_met_d(name = "Isfahan2") + 
  guides(fill = guide_legend(override.aes = list(size = .5), ncol = 2, byrow = FALSE))



# Number of DEGs by sorting status
Fig5j <- ggvenn(data = DEGs_Sorting, 
                show_percentage = FALSE, 
                set_name_size = 2.5, text_size = 2.2,
                fill_color = unlist(as.list(met.brewer(name = "Egypt", n = 3))), 
                stroke_alpha = 0, 
                stroke_color = "white", 
                show_elements = FALSE)

# Heatmap of top 10 DEGs by sorting status (unique DEGs across 3 comparisons)
Fig5k <- DoHeatmap(object = iDA_basal_neu, 
                   #slot = "scale.data",
                   features = DEGs_Sorting_Top10, 
                   lines.width = 10,
                   group.by = "BFP", 
                   group.bar.height = .05,  
                   angle = 0, 
                   size = 4,
                   group.colors = unlist(as.list(met.brewer(name = "Isfahan2", n = 3)))) + 
  theme(legend.title = element_blank(),
        legend.position = "bottom", 
        legend.direction = "horizontal", 
        legend.box.just = "right", 
        axis.text.y = element_text(size = 8))

Fig5k$data$Feature <- gene_info$Gene_Name[match(x = Fig5k$data$Feature, table = gene_info$Gene_ID)]

```

```{r fig5pdf}

pdf(file = "./Figures/Fig5_v3.pdf", height = 6, width = 7)
plot_grid(plot_grid(Fig5f, Fig5g, Fig5h, Fig5i, 
                    ncol = 4, 
                    align = "h", 
                    axis = "bt", 
                    labels = c("F", "G", "H", "I"),
                    rel_widths = c(1, 1.1, 1, 1)),
          plot_grid(Fig5j, Fig5k, 
                    ncol = 2, 
                    align = "h", 
                    axis = "bt", 
                    labels = c("J", "K"), 
                    rel_widths = c(1, 1.7)), 
          ncol = 1, rel_heights = c(1, 1.2))
dev.off()


# Save as Objects
saveRDS(object = Fig5f, "./Figures/Fig5f.rds")
saveRDS(object = Fig5g, "./Figures/Fig5g.rds")
saveRDS(object = Fig5h, "./Figures/Fig5h.rds")
saveRDS(object = Fig5i, "./Figures/Fig5i.rds")
saveRDS(object = Fig5j, "./Figures/Fig5j.rds")
saveRDS(object = Fig5k, "./Figures/Fig5k.rds")

rm(Fig5f, Fig5g, Fig5h, Fig5i, Fig5j, Fig5k)

```

```{r figS4}

# A ) Expression levels of tagBFP
FigS4a <- VlnPlot(object = iDA_basal, features = "ENSG00000000000", group.by = "BFP", assay = "RNA") + 
  theme_cowplot() + 
  scale_fill_met_d(name = "Isfahan2") + 
  theme(title = element_text(hjust = .5),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 8),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.key.size = unit(x = .3, units = "cm"),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
        legend.spacing.x = unit(.5, "mm"),
        legend.spacing.y = unit(2, "mm"),
        legend.text = element_text(size = 8),
        legend.position = "top", 
        legend.justification = "center", 
        legend.direction = "horizontal", 
        legend.title = element_blank()) +
  ggtitle(label = "")

# B) Number of differentially expressed genes per cluster (Cluster markers basal)
FigS4b <- ggplot(data = nDEG_Cluster_basal, mapping = aes(x = Cluster, y = DEG, fill = Cluster)) + 
  geom_col() + 
  scale_fill_d3(palette = "category10") + 
  theme_cowplot() +
  theme(axis.title = element_text(size = 10),
        axis.text = element_text(size = 8),
        legend.key.size = unit(x = .3, units = "cm"),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
        legend.spacing.x = unit(.5, "mm"),
        legend.spacing.y = unit(2, "mm"),
        legend.text = element_text(size = 8),
        legend.title = element_blank(),
        legend.position = "top", 
        legend.justification = "center", 
        legend.direction = "horizontal") + 
  ylab("DEGs by Cluster") + 
  geom_text(aes(label = DEG)) +
  guides(fill = guide_legend(override.aes = list(size = .5), ncol = 4))


# C) Top gene markers per Basal cluster
FigS4c <- DotPlot(iDA_basal, features = DEGs_Cluster_top_basal, cluster.idents = TRUE) +
  theme(axis.title = element_text(size = 10),
        axis.text = element_text(size = 8),
        axis.text.y = element_text(size = 6, hjust = 1, vjust = .5), 
        legend.key.size = unit(x = .3, units = "cm"),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
        legend.spacing.x = unit(.5, "mm"),
        legend.spacing.y = unit(2, "mm"),
        legend.title = element_blank(), 
        legend.position = "top", 
        legend.direction = "horizontal", 
        legend.box = "vertical") +
  coord_flip()
FigS4c$data$features.plot <- factor(gene_info$Gene_Name[match(x = DEGs_Cluster_top_basal, table = gene_info$Gene_ID)], 
                                    levels = rev(gene_info$Gene_Name[match(x = DEGs_Cluster_top_basal, table = gene_info$Gene_ID)]))


# D) All cell type markers from list 
FigS4d <- DotPlot(iDA_basal, features = gm$Gene_ID, cluster.idents = TRUE) +
  theme(axis.title = element_text(size = 10),
        axis.text = element_text(size = 8),
        axis.text.y = element_text(size = 6, hjust = 1, vjust = .5), 
        legend.key.size = unit(x = .3, units = "cm"),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
        legend.spacing.x = unit(.5, "mm"),
        legend.spacing.y = unit(2, "mm"),
        legend.title = element_blank(), 
        legend.position = "top", 
        legend.direction = "horizontal", 
        legend.box = "vertical") +
  coord_flip()
FigS4d$data$features.plot <- factor(gene_info$Gene_Name[match(x = gm$Gene_ID, 
                                                              table = gene_info$Gene_ID)], 
                                    levels = rev(gene_info$Gene_Name[match(x = gm$Gene_ID, table = gene_info$Gene_ID)]))

```

```{r figS4pdf}


pdf("./Figures/FigS4_v3.pdf", height = 9, width = 7)
plot_grid(plot_grid(FigS4a, FigS4b, FigS4c, ncol = 1, labels = c("A", "B", "C"), axis = "lr", align = "v", rel_heights = c(.8, .8, 2)),
          FigS4d, ncol = 2, labels = c("", "D"))

dev.off()

# Save as Objects
saveRDS(object = FigS4a, "./Figures/FigS4a.rds")
saveRDS(object = FigS4b, "./Figures/FigS4b.rds")
saveRDS(object = FigS4c, "./Figures/FigS4c.rds")
saveRDS(object = FigS4d, "./Figures/FigS4d.rds")

rm(FigS4a, FigS4b, FigS4c, FigS4d)


```

```{r figS5}

a <- readRDS(file = "./MPPsorted/GO_BP_DEAwilcox_UP_PositiveVsNegative_Neuronal.rds")
a <- simplify(a)

b <- readRDS(file = "./MPPsorted/GO_BP_DEAwilcox_DOWN_PositiveVsNegative_Neuronal.rds")
b <- simplify(b)

fc <- DEA_pos_neg$avg_log2FC
names(fc) <- DEA_pos_neg$Gene_Name

FigS5a <- cnetplot(a, foldChange = fc, showCategory = 10)
FigS5b <- cnetplot(b, foldChange = fc, showCategory = 10)

rm(a,b,fc)

```

```{r figS5pdf}

pdf("./Figures/FigS5_v3.pdf", height = 9, width = 7)
plot_grid(FigS5a, FigS5b, ncol = 1, labels = c("A", "B"))
dev.off()

# Save as Objects
saveRDS(object = FigS5a, "./Figures/FigS5a.rds")
saveRDS(object = FigS5b, "./Figures/FigS5b.rds")

rm(FigS5a, FigS5b)

```

```{r fig6}

# D UMAP coloured by Treatment
Fig6d <- ggplot(data = pheno_both, mapping = aes(x = UMAP_1, y = UMAP_2, color = Treatment)) +
  geom_point(size = .3) + 
  scale_color_aaas() + 
  theme_cowplot() + 
  theme(axis.title = element_text(size = 10),
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        legend.key.size = unit(x = .3, units = "cm"),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
        legend.spacing.x = unit(.5, "mm"),
        legend.spacing.y = unit(2, "mm"),
        legend.text = element_text(size = 8),
        legend.position = "top", 
        legend.justification = "center", 
        legend.direction = "horizontal", 
        legend.title = element_blank()) +
  ggtitle(label = "") + 
  xlab(label = "UMAP 1") + 
  ylab(label = "UMAP 2") + 
  guides(colour = guide_legend(override.aes = list(size = 2))) 

# E UMAP coloured by Sorting Status
Fig6e <- ggplot(data = pheno_both, mapping = aes(x = UMAP_1, y = UMAP_2, color = BFP)) +
  geom_point(size = .3) + 
  scale_color_met_d(name = "Isfahan2") + 
  theme_cowplot() + 
  theme(axis.title = element_text(size = 10),
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        legend.key.size = unit(x = .3, units = "cm"),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
        legend.spacing.x = unit(.5, "mm"),
        legend.spacing.y = unit(2, "mm"),
        legend.text = element_text(size = 8),
        legend.position = "top", 
        legend.justification = "center", 
        legend.direction = "horizontal", 
        legend.title = element_blank()) +
  ggtitle(label = "") + 
  xlab(label = "UMAP 1") + 
  ylab(label = "UMAP 2") + 
  guides(colour = guide_legend(override.aes = list(size = 2))) 

# F UMAP coloured by cluster in MPP+ cells
cls <- aggregate.data.frame(x = pheno_MPP[c("UMAP_1", "UMAP_2")], 
                            by = list(pheno_MPP$seurat_clusters), 
                            FUN = median)
cls <- data.frame(cls, Cell_type = pheno_MPP$Cell_type[match(cls$Group.1, 
                                                             table = pheno_MPP$seurat_clusters)])

Fig6f <- ggplot(data = pheno_MPP, mapping = aes(x = UMAP_1, y = UMAP_2, color = seurat_clusters)) +
  geom_point(size = .3) + 
  scale_color_met_d(name = "VanGogh2") + 
  theme_cowplot() + 
  theme(axis.title = element_text(size = 10),
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        legend.key.size = unit(x = .3, units = "cm"),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
        legend.spacing.x = unit(.5, "mm"),
        legend.spacing.y = unit(2, "mm"),
        legend.text = element_text(size = 8),
        legend.position = "top", 
        legend.justification = "center", 
        legend.direction = "horizontal", 
        legend.title = element_blank()) +
  ggtitle(label = "") + 
  xlab(label = "UMAP 1") + 
  ylab(label = "UMAP 2") + 
  guides(colour = guide_legend(override.aes = list(size = 2))) 
Fig6f <- Fig6f + annotate("text", x = cls$UMAP_1, y = cls$UMAP_2, label = cls$Cell_type, size = 2)


# G. UMAP coloured by BFP sorting status
Fig6g <- ggplot(data = pheno_MPP, mapping = aes(x = UMAP_1, y = UMAP_2, color = BFP)) +
  geom_point(size = .3) + 
  scale_color_met_d(name = "Isfahan2") + 
  theme_cowplot() + 
  theme(axis.title = element_text(size = 10),
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        legend.key.size = unit(x = .3, units = "cm"),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
        legend.spacing.x = unit(.5, "mm"),
        legend.spacing.y = unit(2, "mm"),
        legend.text = element_text(size = 8),
        legend.position = "top", 
        legend.justification = "center", 
        legend.direction = "horizontal", 
        legend.title = element_blank()) +
  ggtitle(label = "") + 
  xlab(label = "UMAP 1") + 
  ylab(label = "UMAP 2") + 
  guides(colour = guide_legend(override.aes = list(size = 2))) 

# H A few markers
mkr <- c("HMGB1", "HMGB2", "NR2F1", "ZEB2", "CDH2", "SOX9", "SNAP25", "SYT1", "NEFL", "NEFM", "SST", "MAOA", "OTX2")
Fig6h <- DotPlot(object = iDA_MPP, 
                 features = gene_info$Gene_ID[match(x = mkr, table = gene_info$Gene_Name)], 
                 dot.scale = 3) +
  theme(axis.title = element_text(size = 10), 
        axis.text.y = element_text(size = 8, hjust = 1, vjust = .5),
        axis.text.x = element_text(size = 8),
        legend.key.size = unit(x = .3, units = "cm"),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
        legend.spacing.x = unit(.5, "mm"),
        legend.spacing.y = unit(2, "mm"),
        legend.text = element_text(size = 8),
        legend.title = element_blank(), 
        legend.position = "top", 
        legend.direction = "horizontal",
        legend.box = "vertical") +
  coord_flip()

Fig6h$data$features.plot <- factor(gene_info$Gene_Name[match(x = Fig6h$data$features.plot, 
                                                             table = gene_info$Gene_ID)], 
                                   levels = rev(mkr))


# Include proportion of M cluster by sorting status.
Fig6i <- ggplot(data = pheno_MPP, mapping = aes(x = seurat_clusters, fill = BFP)) + 
  geom_bar() + 
  theme_cowplot() +
  theme(axis.title = element_text(size = 10),
        axis.text = element_text(size = 8),
        legend.key.size = unit(x = .3, units = "cm"),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
        legend.spacing.x = unit(.5, "mm"),
        legend.spacing.y = unit(2, "mm"),
        legend.text = element_text(size = 8),
        legend.title = element_blank(),
        legend.position = "top", 
        legend.justification = "center", 
        legend.direction = "vertical") + 
  ylab("Number of cells") + 
  xlab(label = "Cluster") + 
  scale_fill_met_d(name = "Isfahan2") + 
  guides(fill = guide_legend(override.aes = list(size = .5), ncol = 2, byrow = FALSE))



# I. Proportion of cells in neuronal clusters 
pcc <- melt(prop_clusters[c(1,2, 4), ])
colnames(pcc) <- c("Cluster", "Treatment", "Proportion")

Fig6j <- ggplot(data = pcc, mapping = aes(x = Cluster, fill = Treatment, y = Proportion)) + 
  geom_col(position = "dodge") + 
  theme_cowplot() + 
  scale_fill_aaas() + 
  theme(axis.title = element_text(size = 10),
        axis.text = element_text(size = 8),
        legend.key.size = unit(x = .3, units = "cm"),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
        legend.spacing.x = unit(.5, "mm"),
        legend.spacing.y = unit(2, "mm"),
        legend.text = element_text(size = 8),
        legend.position = "top", 
        legend.justification = "center", 
        legend.direction = "horizontal", 
        legend.title = element_blank())

rm(pcc)

# J Number of DEGs in MPP compared to basal by  sorting status
Fig6k <- ggvenn(data = DEGs_MPP, 
                show_percentage = FALSE, 
                set_name_size = 2.5, 
                text_size = 2.2,
                fill_color = unlist(as.list(met.brewer(name = "Isfahan2", n = 3)[c(3, 2, 1)])), 
                stroke_alpha = 0, 
                stroke_color = "white", 
                show_elements = FALSE)

# Heatmap top DEG in response to PP
Fig6l <- DoHeatmap(object = iDA_both_neu, 
                   features = DEGs_MPP_Top20, 
                   lines.width = 10,
                   group.by = "Treatment", 
                   group.bar.height = .05,  
                   angle = 0, 
                   size = 4, 
                   group.colors = pal_aaas()(2)) + 
  theme(legend.title = element_blank(), legend.position = "bottom", legend.direction = "horizontal", 
        legend.box.just = "right", 
        axis.text.y = element_text(size = 8))

Fig6l$data$Feature <- gene_info$Gene_Name[match(x = Fig6l$data$Feature, table = gene_info$Gene_ID)]


```

```{r fig6pdf}

pdf(file = "./Figures/Fig6_v3.pdf", width = 7, height = 9)
plot_grid(plot_grid(Fig6d, Fig6e, Fig6f, Fig6g, ncol = 4, align = "hv", axis = "bt", labels = c("D", "E", "F", "G")),
          plot_grid(Fig6h, Fig6i, Fig6j, ncol = 3, align = "v", axis = "b", labels = c("H", "I", "J")),
          plot_grid(Fig6k, Fig6l, ncol = 2, labels = c("K", "L"), rel_widths = c(1, 2)), 
          ncol = 1, rel_heights = c(1, 1.1, 1.7))
dev.off()

# Save as Objects
saveRDS(object = Fig6d, "./Figures/Fig5d.rds")
saveRDS(object = Fig6e, "./Figures/Fig5e.rds")
saveRDS(object = Fig6f, "./Figures/Fig5f.rds")
saveRDS(object = Fig6g, "./Figures/Fig5g.rds")
saveRDS(object = Fig6h, "./Figures/Fig5h.rds")
saveRDS(object = Fig6i, "./Figures/Fig5i.rds")
saveRDS(object = Fig6j, "./Figures/Fig5j.rds")
saveRDS(object = Fig6k, "./Figures/Fig5k.rds")

rm(Fig6d, Fig6e, Fig6f, Fig6g, Fig6h, Fig6i, Fig6j, Fig6k, Fig6l)

```

```{r figS6}

# A) Number of differentially expressed genes per cluster (Cluster markers MPP)
FigS6a <- ggplot(data = nDEG_Cluster_MPP, mapping = aes(x = Cluster, y = DEG, fill = Cluster)) + 
  geom_col() + 
  scale_fill_met_d(name = "VanGogh2") + 
  theme_cowplot() +
  theme(axis.title = element_text(size = 10),
        axis.text = element_text(size = 8),
        legend.key.size = unit(x = .3, units = "cm"),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
        legend.spacing.x = unit(.5, "mm"),
        legend.spacing.y = unit(2, "mm"),
        legend.text = element_text(size = 8),
        legend.title = element_blank(),
        legend.position = "top", 
        legend.justification = "center", 
        legend.direction = "horizontal") + 
  ylab("DEGs by Cluster") + 
  geom_text(aes(label = DEG), size = 2.5) +
  guides(fill = guide_legend(override.aes = list(size = .5), ncol = 4))


# B) Top gene markers per MPP cluster
FigS6b <- DotPlot(iDA_MPP, features = DEGs_Cluster_top_MPP, cluster.idents = TRUE) +
  theme(axis.title = element_text(size = 10),
        axis.text = element_text(size = 8),
        axis.text.y = element_text(size = 6, hjust = 1, vjust = .5), 
        legend.key.size = unit(x = .3, units = "cm"),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
        legend.spacing.x = unit(.5, "mm"),
        legend.spacing.y = unit(2, "mm"),
        legend.title = element_blank(), 
        legend.position = "top", 
        legend.direction = "horizontal", 
        legend.box = "vertical") +
  coord_flip()
FigS6b$data$features.plot <- factor(gene_info$Gene_Name[match(x = DEGs_Cluster_top_MPP, table = gene_info$Gene_ID)], 
                                    levels = rev(gene_info$Gene_Name[match(x = DEGs_Cluster_top_MPP, table = gene_info$Gene_ID)]))


# C) Known gene markers 
FigS6c <- DotPlot(iDA_MPP, features = gm$Gene_ID, cluster.idents = TRUE) +
  theme(axis.title = element_text(size = 10), 
        axis.text.x = element_text(size = 8),
        axis.text.y = element_text(size = 6, hjust = 1, vjust = .5), 
        legend.key.size = unit(x = .3, units = "cm"),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
        legend.spacing.x = unit(.5, "mm"),
        legend.spacing.y = unit(2, "mm"),
        legend.title = element_blank(), 
        legend.position = "top", 
        legend.direction = "horizontal", 
        legend.box = "vertical") +
  coord_flip()
FigS6c$data$features.plot <- factor(gene_info$Gene_Name[match(x = gm$Gene_ID, 
                                                              table = gene_info$Gene_ID)], 
                                    levels = rev(gene_info$Gene_Name[match(x = gm$Gene_ID, table = gene_info$Gene_ID)]))


# D) UMAP MPP data labelled with predicted basal clusters
FigS6d <- ggplot(data = pheno_MPP, mapping = aes(x = UMAP_1, y = UMAP_2, color = Predicted_Basal)) +
  geom_point(size = .3) + 
  scale_color_d3() + 
  theme_cowplot() + 
  theme(axis.title = element_text(size = 10),
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        legend.key.size = unit(x = .3, units = "cm"),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
        legend.spacing.x = unit(.5, "mm"),
        legend.spacing.y = unit(2, "mm"),
        legend.text = element_text(size = 8),
        legend.position = "top", 
        legend.justification = "center", 
        legend.direction = "horizontal", 
        legend.title = element_blank()) +
  ggtitle(label = "") + 
  xlab(label = "UMAP 1") + 
  ylab(label = "UMAP 2") + 
  guides(colour = guide_legend(override.aes = list(size = 2))) 

# E) UMAP MPP data prediction score
FigS6e <- ggplot(data = pheno_MPP, mapping = aes(x = UMAP_1, y = UMAP_2, color = Max_Prediction_Score)) +
  geom_point(size = .3) + 
  scale_color_material("indigo") +
  theme_cowplot() + 
  theme(axis.title = element_text(size = 10),
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        legend.key.size = unit(x = .3, units = "cm"),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
        legend.spacing.x = unit(.5, "mm"),
        legend.spacing.y = unit(2, "mm"),
        legend.text = element_text(size = 8),
        legend.position = "top", 
        legend.justification = "center", 
        legend.direction = "horizontal", 
        legend.title = element_blank()) +
  ggtitle(label = "") + 
  xlab(label = "UMAP 1") + 
  ylab(label = "UMAP 2") 

# F) Bar plot with number of cells in basal and predicted basal clusters.
pcc <- data.frame(Treatment = rownames(cell_cluster), cell_cluster)
pcc <- melt(pcc)
colnames(pcc)[2:3] <- c("Cluster", "Cells")

FigS6f <- ggplot(data = pcc, mapping = aes(x = Cluster, fill = Treatment, y = Cells)) + 
  geom_col(position = "dodge") + 
  theme_cowplot() + 
  scale_fill_aaas() + 
  theme(axis.title = element_text(size = 10),
        axis.text = element_text(size = 8),
        legend.key.size = unit(x = .3, units = "cm"),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
        legend.spacing.x = unit(.5, "mm"),
        legend.spacing.y = unit(2, "mm"),
        legend.text = element_text(size = 8),
        legend.position = "top", 
        legend.justification = "center", 
        legend.direction = "horizontal", 
        legend.title = element_blank()) + 
  geom_text(aes(label = Cells), size = 2.5)

```

```{r figS6pdf}

pdf(file = "./Figures/FigS6_v3.pdf", width = 7, height = 9)
plot_grid(plot_grid(FigS6a, FigS6d, FigS6e, FigS6f,  
                    ncol = 1, labels = c("A", "D", "E", "F"), align = "hv", axis = "lrtb"),
          FigS6b, FigS6c, ncol = 3, labels = c("", "B", "C"))
dev.off()

# Save as Objects
saveRDS(object = FigS6a, "./Figures/FigS6a.rds")
saveRDS(object = FigS6b, "./Figures/FigS6b.rds")
saveRDS(object = FigS6c, "./Figures/FigS6c.rds")
saveRDS(object = FigS6d, "./Figures/FigS6d.rds")
saveRDS(object = FigS6e, "./Figures/FigS6e.rds")
saveRDS(object = FigS6f, "./Figures/FigS6f.rds")

rm(FigS6a, FigS6b, FigS6c, FigS6d, FigS6e, FigS6f)

```

# Save analysis

```{r}

# Differential expression contrasting sorting status in neuronal populations
write.table(DEA_pos_neg, file = "./MPPsorted/DEA_pos_neg.csv", sep = ",")
write.table(DEA_pos_NS, file = "./MPPsorted/DEA_pos_NS.csv", sep = ",")
write.table(DEA_neg_NS, file = "./MPPsorted/DEA_neg_NS.csv", sep = ",")


# Differential expression in response to MPP+ per sorted population of neurons
write.table(DEA_MPP_basal_positive, file = "./MPPsorted/DEA_MPP_basal_positive.csv", sep = ",")
write.table(DEA_MPP_basal_negative, file = "./MPPsorted/DEA_MPP_basal_negative.csv", sep = ",")
write.table(DEA_MPP_basal_NS, file = "./MPPsorted/DEA_MPP_basal_NS.csv", sep = ",")


# Cluster markers Basal
write.table(DEA_basal_b0, file = "./MPPsorted/DEA_basal_b0.csv", sep = ",")
write.table(DEA_basal_b1, file = "./MPPsorted/DEA_basal_b1.csv", sep = ",")
write.table(DEA_basal_b2, file = "./MPPsorted/DEA_basal_b2.csv", sep = ",")
write.table(DEA_basal_b3, file = "./MPPsorted/DEA_basal_b3.csv", sep = ",")
write.table(DEA_basal_b4, file = "./MPPsorted/DEA_basal_b4.csv", sep = ",")
write.table(DEA_basal_b5, file = "./MPPsorted/DEA_basal_b5.csv", sep = ",")


# Cluster markers MPP
write.table(DEA_MPP_m0, file = "./MPPsorted/DEA_MPP_m0.csv", sep = ",")
write.table(DEA_MPP_m1, file = "./MPPsorted/DEA_MPP_m1.csv", sep = ",")
write.table(DEA_MPP_m2, file = "./MPPsorted/DEA_MPP_m2.csv", sep = ",")
write.table(DEA_MPP_m3, file = "./MPPsorted/DEA_MPP_m3.csv", sep = ",")
write.table(DEA_MPP_m4, file = "./MPPsorted/DEA_MPP_m4.csv", sep = ",")
write.table(DEA_MPP_m5, file = "./MPPsorted/DEA_MPP_m5.csv", sep = ",")



# Differential expression contrasting sorting status in neuronal populations
saveRDS(DEA_pos_neg, file = "./MPPsorted/DEA_pos_neg.rds")
saveRDS(DEA_pos_NS, file = "./MPPsorted/DEA_pos_NS.rds")
saveRDS(DEA_neg_NS, file = "./MPPsorted/DEA_neg_NS.rds")

# Differential expression in response to MPP+ per sorted population of neurons
saveRDS(DEA_MPP_basal_positive, file = "./MPPsorted/DEA_MPP_basal_positive.rds")
saveRDS(DEA_MPP_basal_negative, file = "./MPPsorted/DEA_MPP_basal_negative.rds")
saveRDS(DEA_MPP_basal_NS, file = "./MPPsorted/DEA_MPP_basal_NS.rds")

# Cluster markers Basal
saveRDS(DEA_basal_b0, file = "./MPPsorted/DEA_basal_b0.rds")
saveRDS(DEA_basal_b1, file = "./MPPsorted/DEA_basal_b1.rds")
saveRDS(DEA_basal_b2, file = "./MPPsorted/DEA_basal_b2.rds")
saveRDS(DEA_basal_b3, file = "./MPPsorted/DEA_basal_b3.rds")
saveRDS(DEA_basal_b4, file = "./MPPsorted/DEA_basal_b4.rds")
saveRDS(DEA_basal_b5, file = "./MPPsorted/DEA_basal_b5.rds")

# Cluster markers MPP
saveRDS(DEA_MPP_m0, file = "./MPPsorted/DEA_MPP_m0.rds")
saveRDS(DEA_MPP_m1, file = "./MPPsorted/DEA_MPP_m1.rds")
saveRDS(DEA_MPP_m2, file = "./MPPsorted/DEA_MPP_m2.rds")
saveRDS(DEA_MPP_m3, file = "./MPPsorted/DEA_MPP_m3.rds")
saveRDS(DEA_MPP_m4, file = "./MPPsorted/DEA_MPP_m4.rds")
saveRDS(DEA_MPP_m5, file = "./MPPsorted/DEA_MPP_m5.rds")

#rm(DEA_basal_b0, DEA_basal_b1, DEA_basal_b2, DEA_basal_b3, DEA_basal_b4, DEA_basal_b5, DEA_basal_b6)
#rm(DEA_MPP_m0, DEA_MPP_m1, DEA_MPP_m2, DEA_MPP_m3, DEA_MPP_m4, DEA_MPP_m5, DEA_MPP_c6)
#rm(DEA_MPP_basal_negative, DEA_MPP_basal_negative, DEA_MPP_basal_NS)
#rm(DEA_pos_neg, DEA_pos_NS, DEA_neg_NS)

```

# Save data


```{r}

saveRDS(iDA, file = "./MPPsorted/iDA.Rds")
saveRDS(iDA_basal, file = "./MPPsorted/iDA_basal.Rds")
saveRDS(iDA_MPP, file = "./MPPsorted/iDA_MPP.Rds")
saveRDS(iDA_basal_neu, file = "./MPPsorted/iDA_basal_neu.Rds")
saveRDS(iDA_both, file = "./MPPsorted/iDA_both.Rds")
saveRDS(iDA_both_neu, file = "./MPPsorted/iDA_both_neu.Rds")

#rm(iDA, iDA_basal, iDA_basal_neu, iDA_MPP, iDA_both, iDA_both_neu)

```


